### 2013-12-04 paper.tex
initial import

\documentclass[twoside]{article}

% ------
% Fonts and typesetting settings
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\linespread{1.05} % Palatino needs more space between lines
\usepackage{microtype}

% ------
% Page layout
\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
\usepackage[font=it]{caption}
\usepackage{paralist}
\usepackage{multicol}

% ------
% Lettrines
\usepackage{lettrine}

% ------
% Abstract
\usepackage{abstract}
	\renewcommand{\abstractnamefont}{\normalfont\bfseries}
	\renewcommand{\abstracttextfont}{\normalfont\small\itshape}

% ------
% Titling (section/subsection)
\usepackage{titlesec}
\renewcommand\thesection{\Roman{section}}
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{}

% ------
% Header/footer
\usepackage{fancyhdr}
	\pagestyle{fancy}
	\fancyhead{}
	\fancyfoot{}
	\fancyhead[C]{Journal paper template $\bullet$ January 2014 $\bullet$ Vol 1}
	\fancyfoot[RO,LE]{\thepage}

% ------
% Clickable URLs (optional)
\usepackage{hyperref}

% ------
% BibTex for bibliography
\usepackage[numbers]{natbib}
%\usepackage[square,sort&compress,authoryear]{natbib}

% ------
% Math
\usepackage{amsthm}
\usepackage{amsmath, mathtools}
\usepackage{fixltx2e}
\newtheorem{mydef}{Definition}
% \setlength{\parskip}{0cm plus4mm minus3mm}

% ------
% Maketitle metadata
\title{\vspace{-15mm}%
	\fontsize{24pt}{10pt}\selectfont
	\textbf{Symmetric NAT traversal algorithm toolbox}
	}	
\author{%
	\large
	\textsc{Du\v{s}an Klinec} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
	\normalsize	Faculty of Informatics \\
	\normalsize	Masaryk University, Brno \\
	\normalsize	\href{mailto:xklinec@mail.muni.cz}{xklinec@mail.muni.cz}
	\and
	\textsc{Vashek Matyáš} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
	\normalsize	Faculty of Informatics \\
	\normalsize	Masaryk University, Brno \\
	\normalsize	\href{mailto:matyas@mail.muni.cz}{matyas@mail.muni.cz}
	\vspace{-5mm}
	}
\date{}

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\maketitle
\thispagestyle{fancy}

\begin{abstract}
\noindent Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur magna lorem, tempor sed facilisis vel, porta et turpis. Sed et felis a massa dictum posuere. Aliquam hendrerit rhoncus ipsum sit amet placerat. Duis fringilla est eu arcu mollis faucibus non sit amet eros. Vestibulum risus nibh, dapibus vitae laoreet eget, fringilla quis nisl. Proin consequat nibh sit amet mauris suscipit tincidunt. Sed rutrum, purus nec aliquam faucibus, quam libero venenatis nisi, ut tempor mi sapien vel diam. Pellentesque sagittis elit non risus malesuada accumsan. Morbi consequat urna et lacus hendrerit sodales. Proin at urna neque, ut dapibus urna. Curabitur venenatis molestie convallis. Vestibulum blandit vulputate risus, quis sodales sapien porttitor non.
\end{abstract}

\begin{multicols}{2}
% \lettrine[nindent=0em,lines=3]{L}orem ipsum dolor sit amet.

\section{Introduction}
Network Addres Translators (abbreviated as NAT) has been increasing in popularity in recent years 
mainly due to IPv4 address space depletion, in order to slow it down. On one side NAT offers
the ability to share a public, globaly routable IP address among many internal network hosts but on 
the other side it often breaks basic end-to-end principle of the Internet and thus favors server-client 
communication model for clients connected behind NAT.

In many applications is peer-to-peer (P2P) connection of crucial importance and the quality of service depends on 
its quality heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
between communicating parties is necessary. If P2P connection is not possible to establish relay 
servers has to be used (TURN~\citep{rfc5766} protocol is used). The presence 
of another node in communication path can have negative effects on communication channel quality 
(increased latency, jitter, lower bandwidth), relay server is a potential point of failure, 
has to scale with increasing demands of the network, interesting point for attack, etc.
Thus relay servers require non-negligible resources what makes them expensive for VoIP providers and
establishing direct connection between parties is of high importance.

In this paper we are focused on a symmetric NAT with predictable port allocation function. We propose a 
new algorithm for establishing direct connection between hosts behind such NATs that is effective and 
has low requirements.

\section{Network Address Translation}

Hola hop, hola hop, he-ja he-ja hej rup. Hola hop, hola hop, he-ja he-ja hej rup.
\begin{compactitem}
\item STUN~\citep{rfc5389}
\item TURN~\citep{rfc5766}
\item Teredo~\citep{rfc4380}
\item uPNP~\citep{rfc6970}
\item PCP~\citep{rfc6887}
\item ICE~\citep{rfc5245}
\end{compactitem}

Properties of NAT assumed in our work:
\begin{compactitem}
\item symmetric
\item port-restricted firewall
\item port-sensitive allocation
\item incremental port allocation rule
\end{compactitem}

\section{Related work}
\begin{compactitem}
\item chineese\citep{Wang:2006:RSN:1156422.1156550}
\item uPNP
\item PCP~\citep{rfc6887}
\end{compactitem}

\section{Theoretical network model}
In order to evaluate algorithms in multiple environments with different parameters (workload, number of clients, etc.)
we decided to model network behavior with stochastic process. In particular we choose a Poisson process, $\{N(t), t\geq0\}$ 
that is used in queueing theory to model arrivals of requests to the server queue~\citep{Nelson:1995:PSP:207382}.

The core idea of NAT traversal is to predict the next allocated external port by NAT, thus from this perspective 
it is important to model how NAT's internal state changes over time, i.e. how many external ports 
were allocated between two time intervals. Thus the workload is important parameter for predicting allocated port.

We model internal network connected to NAT w.r.t. newly created connections (new port allocation) as a 
time homogenous Poisson process $Po(\lambda)$. By setting $\lambda$ we can model 
different workload of the network and evaluate algorithms with respect to the various parameters e.g., 
probability of an establishing connection.

% Random variable $X_t \sim Po(\lambda t)$ says how many
% port allocations were made in time interval $[0,t]$

\begin{mydef}
$N(t)$ is a random variable of a number of a new port allocations made in time 
interval~$[0,t]$ where $N(t) \sim Po(\lambda t)$. \\

\begin{center}                                                     
Then $P[N(t)=n] = \frac{(\lambda t)^n}{n!} e^{-\lambda t}$
\end{center}
\end{mydef}

\section{UDP hole punching}
All algorithms here make use of the same principle, UDP hole punching. NAT separates internal
IP address space from the outer world, thus internet hosts cannot send packet to the 
internal host directly. Communication has to be initiated from the inside\footnote{Ignoring 
pre-set port forwarding, DMZ, uPnP, etc\dots}.

 By default a packet 
comming from the internet (\textit{IP\textsubscript{A},Port\textsubscript{A}}) to the internal 
network is dropped on the NAT unless there is a valid mapping in NAT allocation table.

\section{Generic algorithm structure}
Algorithms we propose have a common structure. In order to maximize probability of an establishing
a connection (i.e., success rate) it proceeds in discrete \emph{steps}. In each step it 
tries to open a new hole on it's NAT and/or to predict corresponding port allocated on the destination
NAT.

\begin{mydef}
Symmetric NAT traversal algorithm is a mapping 
$\mathcal{A}: \mathcal{U} \times \mathcal{S} \rightarrow \mathcal{P}_{src} \times \mathcal{P}_{dst}$, 
where \\
\begin{compactitem}
\item $\mathcal{U}=\{0,1\}$ is set of communicating parties
\item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$
is a step of the algorithm
\item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
are sets of source and destination ports respectively
\end{compactitem}
\end{mydef}

\paragraph{Intuition.} 
If we have $A(0,1) = (2,3)$ it means party $0$ in step $1$ sent packet from source port $2$ to destination port $3$
of the party $1$.

\paragraph{Time.} 
The algorithm proceeds in discrete time steps. Algorith starts in time $t_0 = 0$~ms and each step $i \in \mathcal{S}$
happens in time $t_i = iT$ for some $T$. If it is not mentioned otherwise, $T = 10$~ms.

\paragraph{Party change.} 
We define unary operation on communicating parties. 
\[
a \in \mathcal{U},  \neg a = \begin{dcases*}
         0 & when $a=1$\\
         1 & when $a=0$
        \end{dcases*}
\]

\subsection{Algorithm I - baby-step, giant-step}
We propose the algorithm with high success rate for small network loads, i.e.,
$\lambda \leq 0.03, T=10$~ms. When error becomes larger,
algorithm step is not fast enought to compensate errors and thus fails for larger $\lambda$.

\begin{mydef}
Baby-step giant-step algorithm
\[
u \in \mathcal{U}, s \in \mathcal{S}: 
\mathcal{A}(u, s) = \begin{dcases*}
         (0, s)  & when $u=0$\\
         (0, 2s) & when $u=1$
        \end{dcases*}
\]
\end{mydef}

Intuitively the source port is kept constant, one side of the protocol scans ports sequentially, another
side of the protocol scans ports with step 2.

\subsection{Algorithm II - fixed destination port}
Algorithm with fixed destination port described in by \citep{Wang:2006:RSN:1156422.1156550}
changes the source port in order to create new holes in NAT in each step.

\begin{mydef}
Fixed destination port algorithm
\[
u \in \mathcal{U}, s \in \mathcal{S}: \;
\mathcal{A}(u, s) = (s, \Delta_u)
\] for some constants $\Delta_0, \Delta_1$ based on the current network workload determined in initial phase
of the protocol.
\end{mydef}

Intuitively the source port is kept constant, one side of the protocol scans ports sequentially, another
side of the protocol scans ports with step 2.

% -----
% Bibliography
\bibliographystyle{unsrtnat}
\bibliography{paper}

\end{multicols}

\end{document}

### 2013-12-05 paper.tex
paper - few fixes

@@
 We model internal network connected to NAT w.r.t. newly created connections (new port allocation) as a 
-time homogenous Poisson process $Po(\lambda)$. By setting $\lambda$ we can model 
+time homogeneous Poisson process $Po(\lambda)$. By setting $\lambda$ we can model 
 different workload of the network and evaluate algorithms with respect to the various parameters e.g., 
@@
  By default a packet 
-comming from the internet (\textit{IP\textsubscript{A},Port\textsubscript{A}}) to the internal 
+coming from the internet (\textit{IP\textsubscript{A},Port\textsubscript{A}}) to the internal 
 network is dropped on the NAT unless there is a valid mapping in NAT allocation table.
@@
 Symmetric NAT traversal algorithm is a mapping 
-$\mathcal{A}: \mathcal{U} \times \mathcal{S} \rightarrow \mathcal{P}_{src} \times \mathcal{P}_{dst}$, 
+$\mathcal{A}: \mathcal{U} \times \mathcal{S} \times \mathcal{M} \rightarrow \mathcal{P}_{src} \times \mathcal{P}_{dst}$, 
 where \\
@@
 \item $\mathcal{U}=\{0,1\}$ is set of communicating parties
-\item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$
-is a step of the algorithm
+\item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$ is a step of the algorithm
+\item $\mathcal{M}$ contains vectors of parameters of the environment (i.e., network model 
+parameters and additional information obtained in initial phase)
 \item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
@@
 \paragraph{Time.} 
-The algorithm proceeds in discrete time steps. Algorith starts in time $t_0 = 0$~ms and each step $i \in \mathcal{S}$
+The algorithm proceeds in discrete time steps. Algorithm starts in time $t_0 = 0$~ms and each step $i \in \mathcal{S}$
 happens in time $t_i = iT$ for some $T$. If it is not mentioned otherwise, $T = 10$~ms.
@@
 $\lambda \leq 0.03, T=10$~ms. When error becomes larger,
-algorithm step is not fast enought to compensate errors and thus fails for larger $\lambda$.
+algorithm step is not fast enough to compensate errors and thus fails for larger $\lambda$.
 
@@
 
+\subsection{Algorithm III - expected port value}
+AHA!
+
+\subsection{Algorithm IV - Poisson sampling}
+Uhmmmmmmm....
+
 % -----

### 2013-12-05 paper.tex
NAT is formall now

@@
 \section{Introduction}
-Network Addres Translators (abbreviated as NAT) has been increasing in popularity in recent years 
+Network Address Translators (abbreviated as NAT) has been increasing in popularity in recent years 
 mainly due to IPv4 address space depletion, in order to slow it down. On one side NAT offers
@@
 \section{Network Address Translation}
+Common NAT categorization is according to~\citep{rfc3489}: a) Full Cone
+b) IP Restricted Cone, c) Port Restricted Cone, d) Symmetric.
 
-Hola hop, hola hop, he-ja he-ja hej rup. Hola hop, hola hop, he-ja he-ja hej rup.
-\begin{compactitem}
-\item STUN~\citep{rfc5389}
-\item TURN~\citep{rfc5766}
-\item Teredo~\citep{rfc4380}
-\item uPNP~\citep{rfc6970}
-\item PCP~\citep{rfc6887}
-\item ICE~\citep{rfc5245}
-\end{compactitem}
+NAT connects two networks, denoted internal and external in further text. We assume NAT has only 
+one external address for simplicity.
 
-Properties of NAT assumed in our work:
-\begin{compactitem}
-\item symmetric
-\item port-restricted firewall
-\item port-sensitive allocation
-\item incremental port allocation rule
-\end{compactitem}
+\paragraph{Definitions.}
+Symmetric NAT could be defined as follows.
+
+\begin{mydef}
+\begin{align*}
+\mathbb{SC} = \mathbb{IP} \times \mathbb{PORT}
+\end{align*} is a Socket. $\mathbb{IP}$ is set of all possible IP addresses and $\mathbb{PORT}$ 
+is set of all possible ports, basically $\mathbb{PORT} = \{1,2,\dots,65535\}$.
+\begin{align*}
+\mathbb{SP} = \mathbb{IP} \times \mathbb{PORT} \times \mathbb{IP} \times \mathbb{PORT}
+\end{align*} is a Socket Pair.
+\end{mydef}
+
+\begin{mydef}
+\begin{align*}
+\mathbb{AT} \subseteq \mathbb{SP} \times \mathbb{PORT}
+\end{align*} is an allocation table of the NAT. Second component is external port allocated for particular socket pair.
+\end{mydef}
+Allocation table is a set of all socket pairs and corresponding allocated external port.
+
+% \paragraph{Filtering rule.} Let assume packet with socket pair {(IP\textsubscript{out},PORT\textsubscript{out},IP\textsubscript{ex},PORT\textsubscript{ex})} 
+% received on NAT external interface (we'll use this notation in further text). Packet is allowed to pass to internal network iff 
+% \begin{align*}
+% &  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
+% & ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}), \text{PORT\textsubscript{ex}} ) \in \mathbb{AT}
+% \end{align*}
+% 
+% If packet is allowed to pass, it is then forwarded to host {(IP\textsubscript{in}, PORT\textsubscript{in})}. A new record 
+% to the allocation table is added when internal host sends packet to the external host. 
+% 
+% \paragraph{Allocation rule.} ToDo
+
+\subsection{NAT classes}
+
+\paragraph{Full cone NAT.} ~\\
+Allocation table $\mathbb{AT} \subseteq \mathbb{SC} \times \mathbb{PORT}$ holds association between internal 
+hosts socket and mapped external NAT port. 
+% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
+Filtering rule is
+\begin{align*}
+& PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}} )) \Leftrightarrow \\
+&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
+& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}), \text{PORT\textsubscript{ex}} ) \in \mathbb{AT}
+\end{align*}
+
+\paragraph{Address restricted cone NAT.} ~\\
+Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP} \times \mathbb{PORT} \times \mathbb{IP}\} \times \mathbb{PORT}$ 
+holds association between internal hosts socket, external host IP address and mapped external NAT port. 
+% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
+Filtering rule is
+\begin{align*}
+& PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}} )) \Leftrightarrow \\
+&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
+& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{out}}), \text{PORT\textsubscript{ex}} ) \in \mathbb{AT}
+\end{align*}
+
+\paragraph{Port restricted cone NAT.} ~\\
+Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP} \times \mathbb{PORT} \times \mathbb{IP} \times \mathbb{PORT}\} \times \mathbb{PORT}$ 
+holds association between internal hosts socket, external host socket and mapped external NAT port. 
+% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
+Filtering rule is
+\begin{align*}
+& PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}} )) \Leftrightarrow \\
+&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
+& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}), \text{PORT\textsubscript{ex}} ) \in \mathbb{AT}
+\end{align*}
+
+\paragraph{Symmetric NAT.} ~\\
+Allocation table $\mathbb{AT}$ and filtering rule is the same as in Port restricted cone NAT. The difference lies in the way 
+the new entries in allocation table are created. For Cone NAT holds that multiple entries in the allocation table 
+can have the same external port assigned, while Symmetric NAT creates new allocation entry for each new connection. 
+More formally the restriction for $\mathbb{AT}$: 
+\begin{align*}
+& \exists s \in \mathbb{SP}, p \in \mathbb{PORT}: (s, p) \in \mathbb{AT} \Rightarrow (\forall s^{\prime} \in \mathbb{SP} \\
+& s^{\prime} \neq s , p^{\prime} \in \mathbb{PORT}: ((s^{\prime}, p^{\prime}) \in \mathbb{AT}): p \neq p^{\prime})
+\end{align*}
+% 
+%     PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}} )) \Leftrightarrow \\
+% &  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
+% & ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}), \text{PORT\textsubscript{ex}} ) \in \mathbb{AT}
+% \end{align*}
+
+% 
+%We mainly categorize NATs from two perspectives a) 
+% 
+% Hola hop, hola hop, he-ja he-ja hej rup. Hola hop, hola hop, he-ja he-ja hej rup.
+% \begin{compactitem}
+% \item STUN~\citep{rfc5389}
+% \item TURN~\citep{rfc5766}
+% \item Teredo~\citep{rfc4380}
+% \item uPNP~\citep{rfc6970}
+% \item PCP~\citep{rfc6887}
+% \item ICE~\citep{rfc5245}
+% \end{compactitem}
+%
+
+% Properties of NAT assumed in our work:
+% \begin{compactitem}
+% \item symmetric
+% \item port-restricted firewall
+% \item port-sensitive allocation
+% \item incremental port allocation rule
+% \end{compactitem}
+
+\section{Existing techniques}
+Many different solutions for traversing NATs exist e.g., NAT traversal in Internet Key Exchange~\citep{rfc3947} or Terredo tunneling~\citep{rfc4380}.
+
+Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}. But in general 
+it is not widely deployed and it cannot be relied upon.
+
+Most of the applications nowadays use Interactive Connectivity Establishment~(ICE)~\citep{rfc5245} what is a very powerful toolbox for traversing 
+various types of NAT, but does not work with symmetric NATs. ICE mainly uses STUN and TURN.
 
 \section{Related work}
+Some interesting algorithms for traversing symmetric NAT were proposed by Y. Takeda in \citep{takeda}, but the case both clients are behind symmetric NAT
+with port sensitive allocation function is not covered. It mentions that sending more packets won't increase probability of establishing the new connection.
+Our motivation is to show that it is actually possible.
+
+The \citep{rfc5128} maps current state of the art in NAT traversal. But does not cover our case.
+
+Also the academic research was done in this field, for example by Wei~\emph{et.~al.} in \citep{wei}., but their algorithm is using another technique compared
+to our approach. The downside of their algorithm is it requires Time To Live~(TTL) modification of the IP~packet. This is in majority of cases considered
+as a privileged operation in operating system and sufficient privileges are required for this change.
+
+\section{Related work II}
 \begin{compactitem}
@@
 \end{compactitem}
+
+\section{UDP hole punching with STUN}
+All algorithms here make use of the same principle, UDP hole punching. NAT separates internal
+IP address space from the outer world, thus internet hosts cannot send packet to the 
+internal host directly. Communication has to be initiated from the inside\footnote{Ignoring 
+pre-set port forwarding, DMZ, uPnP, etc\dots}, what is problematic if two nodes behind 
+NAT wants to communicate.
+
+By default a packet 
+coming from the internet (\textit{IP\textsubscript{A},Port\textsubscript{A}}) to the internal 
+network is dropped on the NAT unless there is a valid mapping in NAT allocation table.
 
@@
 
-\section{UDP hole punching}
-All algorithms here make use of the same principle, UDP hole punching. NAT separates internal
-IP address space from the outer world, thus internet hosts cannot send packet to the 
-internal host directly. Communication has to be initiated from the inside\footnote{Ignoring 
-pre-set port forwarding, DMZ, uPnP, etc\dots}.
-
- By default a packet 
-coming from the internet (\textit{IP\textsubscript{A},Port\textsubscript{A}}) to the internal 
-network is dropped on the NAT unless there is a valid mapping in NAT allocation table.
-
 \section{Generic algorithm structure}
@@
 \item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$ is a step of the algorithm
-\item $\mathcal{M}$ contains vectors of parameters of the environment (i.e., network model 
-parameters and additional information obtained in initial phase)
+\item $\mathcal{M}$ is set of vectors of parameters of the environment (i.e., network model 
+parameters and additional information obtained in initial phase). 
+In our setting $\mathcal{M} = \{ (\lambda) | \lambda > 0 \}$. For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
 \item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
@@
 \end{mydef}
+If third parameter is not important for the algorithm it can be ignored in notation.
 
@@
 \subsection{Algorithm III - expected port value}
-AHA!
+Some more effective algorithms can be designed if the random process the network follows is known (and their
+parameters).
+
+\begin{mydef}
+Expected port value algorithm
+\[
+u \in \mathcal{U}, s \in \mathcal{S}, \lambda \in \mathcal{M}:\;
+\mathcal{A}(u, s, \lambda) = (0, s(1 + \lambda T))
+\]
+\end{mydef}
 
 \subsection{Algorithm IV - Poisson sampling}
-Uhmmmmmmm....
+Algorithm with the best success rate in our simulation is based on sampling Poisson distribution.
+The main idea is to let the algorithm simulate the Poisson process that the network of the
+other party follows.
+
+\begin{mydef}
+Poisson sampling algorithm 
+\begin{align*}
+& u \in \mathcal{U}, s \in \mathcal{S}, \lambda \in \mathcal{M}:\\
+& \mathcal{A}(u, s, \lambda) = (0, \text{Poisson}(\lambda T (1+s C(\lambda)))
+\end{align*} where $\text{Poisson}(\mu) = X \sim Po(\mu)$ samples Poisson distribution with
+parameter $\mu$, function $C(\lambda)$ gives a constant 
+coefficient\footnote{This function was found empirically by maximizing success rate}.
+\end{mydef}
 

### 2013-12-06 paper.tex
notation and theory

@@
 	\fancyfoot{}
-	\fancyhead[C]{Journal paper template $\bullet$ January 2014 $\bullet$ Vol 1}
+	\fancyhead[C]{Security Protocol Workshop $\bullet$ March 2014 $\bullet$ Vol 1}
 	\fancyfoot[RO,LE]{\thepage}
@@
 \newtheorem{mydef}{Definition}
+\newcommand{\ignore}[1]{}
 % \setlength{\parskip}{0cm plus4mm minus3mm}
@@
 	\fontsize{24pt}{10pt}\selectfont
-	\textbf{Symmetric NAT traversal algorithm toolbox}
+	\textbf{Traversing Symmetric NAT with predictable port allocation function}
 	}	
@@
 \begin{abstract}
-\noindent Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+\noindent ToDo; Will be written as the last part of the paper...
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 Curabitur magna lorem, tempor sed facilisis vel, porta et turpis.
@@
 Pellentesque sagittis elit non risus malesuada accumsan.
-Morbi consequat urna et lacus hendrerit sodales.
-Proin at urna neque, ut dapibus urna.
-Curabitur venenatis molestie convallis.
-Vestibulum blandit vulputate risus, quis sodales sapien porttitor non.
+Morbi consequat urna.
 \end{abstract}
@@
 mainly due to IPv4 address space depletion, in order to slow it down. On one side NAT offers
-the ability to share a public, globaly routable IP address among many internal network hosts but on 
+the ability to share a public, globally routable IP address among many internal network hosts but on 
 the other side it often breaks basic end-to-end principle of the Internet and thus favors server-client 
@@
 
-In many applications is peer-to-peer (P2P) connection of crucial importance and the quality of service depends on 
-its quality heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
-between communicating parties is necessary. If P2P connection is not possible to establish relay 
+In many applications is peer-to-peer (P2P) connection of crucial importance and the quality of a service depends on 
+it's parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
+between communicating parties is necessary. If P2P connection is not possible to be established relay 
 servers has to be used (TURN~\citep{rfc5766} protocol is used). The presence 
-of another node in communication path can have negative effects on communication channel quality 
+of another node in the communication path can have negative effects on a communication channel quality 
 (increased latency, jitter, lower bandwidth), relay server is a potential point of failure, 
@@
 Thus relay servers require non-negligible resources what makes them expensive for VoIP providers and
-establishing direct connection between parties is of high importance.
+establishing a direct connection between parties is of high importance.
 
-In this paper we are focused on a symmetric NAT with predictable port allocation function. We propose a 
-new algorithm for establishing direct connection between hosts behind such NATs that is effective and 
-has low requirements.
+In this paper we are focused on a symmetric NAT with a predictable port allocation function. We propose 
+new algorithms for establishing direct connection between hosts behind such NATs that are effective and 
+has low system requirements.
 
 \section{Network Address Translation}
-Common NAT categorization is according to~\citep{rfc3489}: a) Full Cone
-b) IP Restricted Cone, c) Port Restricted Cone, d) Symmetric.
+The notation we will be using for the rest of the paper follows.
 
-NAT connects two networks, denoted internal and external in further text. We assume NAT has only 
-one external address for simplicity.
-
-\paragraph{Definitions.}
-Symmetric NAT could be defined as follows.
+% \paragraph{Notation.}
+% Symmetric NAT could be defined as follows.
 
@@
 \end{mydef}
+% 
+% \begin{mydef}
+% \begin{align*}
+% \mathbb{AT} \subseteq \mathbb{SP} \times \mathbb{PORT}
+% \end{align*} is an allocation table of the NAT. Second component is external port allocated for particular socket pair.
+% \end{mydef}
+% Allocation table is a set of all socket pairs and corresponding allocated external port.
 
-\begin{mydef}
-\begin{align*}
-\mathbb{AT} \subseteq \mathbb{SP} \times \mathbb{PORT}
-\end{align*} is an allocation table of the NAT. Second component is external port allocated for particular socket pair.
-\end{mydef}
-Allocation table is a set of all socket pairs and corresponding allocated external port.
+\paragraph{NAT and communication.}
+NAT connects two networks, denoted as an internal and an external, connected to an internal and an external NAT interface respectively.
+Hosts has addresses $\text{IP\textsubscript{in}} \in \mathbb{IP}_{in} \subseteq \mathbb{IP}$ 
+and $\text{IP\textsubscript{ex}} \in \mathbb{IP}_{ex} \subseteq \mathbb{IP}$ in the internal and the external network
+respectively, $\mathbb{IP}_{in} \cap \mathbb{IP}_{ex} = \emptyset$.
+If it is not said otherwise the external NAT address\footnote{We assume NAT has only one external address for simplicity.} is denoted as 
+$\text{IP\textsubscript{nat}} \in \mathbb{IP}_{ex}$.
+
+Packet received on the external interface has socket pair {(IP\textsubscript{ex},PORT\textsubscript{ex},IP\textsubscript{nat},PORT\textsubscript{nat})}.
+
+Packet received on the internal interface has socket pair {(IP\textsubscript{in},PORT\textsubscript{in},IP\textsubscript{ex},PORT\textsubscript{ex})}.
 
@@
 \subsection{NAT classes}
+NAT is classified w.r.t. \emph{filtering rule} and \emph{allocation rule} which says what happens after receiving
+packet from an external and an internal network respectively.
+
+Common NAT categorization is according to~\citep{rfc3489}: a) Full Cone
+b) IP Restricted Cone, c) Port Restricted Cone, d) Symmetric.
 
 \paragraph{Full cone NAT.} ~\\
-Allocation table $\mathbb{AT} \subseteq \mathbb{SC} \times \mathbb{PORT}$ holds association between internal 
+Allocation table $\mathbb{AT} \subseteq \{ \mathbb{IP}_{in} \times \mathbb{PORT} \} \times \mathbb{PORT}$ holds associations between internal 
 hosts socket and mapped external NAT port. 
@@
 \begin{align*}
-& PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}} )) \Leftrightarrow \\
-&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
-& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}), \text{PORT\textsubscript{ex}} ) \in \mathbb{AT}
+& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
+&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}_{in}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
+& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
 \end{align*}
@@
 \paragraph{Address restricted cone NAT.} ~\\
-Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP} \times \mathbb{PORT} \times \mathbb{IP}\} \times \mathbb{PORT}$ 
-holds association between internal hosts socket, external host IP address and mapped external NAT port. 
+Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex}\} \times \mathbb{PORT}$ 
+holds associations between internal hosts socket, external host IP address and mapped external NAT port. 
 % Rule is created once the connection from the internal host is initiated to an arbitrary external address.
@@
 \begin{align*}
-& PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}} )) \Leftrightarrow \\
+& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
 &  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
-& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{out}}), \text{PORT\textsubscript{ex}} ) \in \mathbb{AT}
+& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
 \end{align*}
@@
 \paragraph{Port restricted cone NAT.} ~\\
-Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP} \times \mathbb{PORT} \times \mathbb{IP} \times \mathbb{PORT}\} \times \mathbb{PORT}$ 
-holds association between internal hosts socket, external host socket and mapped external NAT port. 
+Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex} \times \mathbb{PORT}\} \times \mathbb{PORT}$ 
+holds associations between internal hosts socket, external host socket and mapped external NAT port. 
 % Rule is created once the connection from the internal host is initiated to an arbitrary external address.
@@
 \begin{align*}
-& PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}} )) \Leftrightarrow \\
+& PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
 &  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
-& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}), \text{PORT\textsubscript{ex}} ) \in \mathbb{AT}
+& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
 \end{align*}
@@
 \paragraph{Symmetric NAT.} ~\\
-Allocation table $\mathbb{AT}$ and filtering rule is the same as in Port restricted cone NAT. The difference lies in the way 
-the new entries in allocation table are created. For Cone NAT holds that multiple entries in the allocation table 
-can have the same external port assigned, while Symmetric NAT creates new allocation entry for each new connection. 
-More formally the restriction for $\mathbb{AT}$: 
+Allocation table $\mathbb{AT}$ and filtering rule is the same as in Port restricted Cone NAT. The difference lies in the way 
+the new entries in the allocation table are created. For Cone NAT holds that multiple entries in the allocation table 
+can have the same external port assigned, while Symmetric NAT creates new an allocation entries under some circumstances.
+
+\emph{Allocation function (AF)} says how is the port allocated for new connection chosen. Assume $\mathbb{AT}=\emptyset$, for simplicity. 
+AF can be \emph{predictable}, e.g. \emph{incremental}. If previously allocated port was $\text{PORT\textsubscript{nat}}$
+the next port will be $\text{PORT\textsubscript{nat}}+\Delta$, where $\Delta=1$ usually. Another type of AF is \emph{random}.
+If the chosen port is already taken, AF iterates until the free one is found.
+
+AF can have different \emph{sensitivity} determining the cases where new allocation is created or the previous one is used instead.
+
+\emph{Address sensitive allocation function} creates new allocation if a destination address differs from the existing allocation 
+from the same source:
 \begin{align*}
-& \exists s \in \mathbb{SP}, p \in \mathbb{PORT}: (s, p) \in \mathbb{AT} \Rightarrow (\forall s^{\prime} \in \mathbb{SP} \\
-& s^{\prime} \neq s , p^{\prime} \in \mathbb{PORT}: ((s^{\prime}, p^{\prime}) \in \mathbb{AT}): p \neq p^{\prime})
+& \exists s=(a_1,a_2,a_3,a_4) \in \mathbb{SP}, p \in \mathbb{PORT}: \\
+& (s, p) \in \mathbb{AT} \Rightarrow ((\forall s^{\prime}=(a_1^{\prime},a_2^{\prime},a_3^{\prime},a_4^{\prime}) \in \mathbb{SP}, \\
+& a_1 \neq a_1^{\prime}, a_2 \neq a_2^{\prime}, a_3 \neq a_3^{\prime}, \forall p^{\prime} \in \mathbb{PORT}: \\
+& ((s^{\prime}, p^{\prime}) \in \mathbb{AT})) \Rightarrow p \neq p^{\prime})
 \end{align*}
+
+\emph{Port sensitive allocation function} creates new allocation if a destination address and a port differs from the existing
+allocations from the same source\footnote{for each new connection is allocated new a NAT port.}:
+\begin{align*}
+& \exists s \in \mathbb{SP}, p \in \mathbb{PORT}: (s, p) \in \mathbb{AT} \Rightarrow \\
+& \Rightarrow ((\forall s^{\prime} \in \mathbb{SP}, s^{\prime} \neq s , \forall p^{\prime} \in \mathbb{PORT}: \\
+& ((s^{\prime}, p^{\prime}) \in \mathbb{AT})) \Rightarrow p \neq p^{\prime})
+\end{align*}
+
 % 
@@
 \section{Existing techniques}
-Many different solutions for traversing NATs exist e.g., NAT traversal in Internet Key Exchange~\citep{rfc3947} or Terredo tunneling~\citep{rfc4380}.
+Many different solutions for traversing non-symmetric NATs exist e.g., NAT traversal in Internet Key Exchange~\citep{rfc3947} or Terredo\ignore{tunneling}~\citep{rfc4380}.
 
@@
 
-Most of the applications nowadays use Interactive Connectivity Establishment~(ICE)~\citep{rfc5245} what is a very powerful toolbox for traversing 
-various types of NAT, but does not work with symmetric NATs. ICE mainly uses STUN and TURN.
+Most of the applications use Interactive Connectivity Establishment~(ICE)~\citep{rfc5245} nowadays, what is a very powerful toolbox for traversing 
+various types of NAT, however does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN~\citep{rfc5766} for it's operation.
 
@@
 to our approach. The downside of their algorithm is it requires Time To Live~(TTL) modification of the IP~packet. This is in majority of cases considered
-as a privileged operation in operating system and sufficient privileges are required for this change.
+as a privileged operation in operating system and sufficient privileges are required for this modification.
 
-\section{Related work II}
-\begin{compactitem}
-\item chineese\citep{Wang:2006:RSN:1156422.1156550}
-\item uPNP
-\item PCP~\citep{rfc6887}
-\end{compactitem}
+In our work we study the situation where two parties, A,B tries to establish a direct connection, both of them are behind 
+a symmetric NAT (possibly cascade of such NATs) with port sensitive, predictable allocation function. Another scenarios with
+predictable allocation function are quite easy to solve and covered in previous research.
+
+% \section{Related work II}
+% \begin{compactitem}
+% \item chineese\citep{Wang:2006:RSN:1156422.1156550}
+% \item uPNP
+% \item PCP~\citep{rfc6887}
+% \end{compactitem}
 
 \section{UDP hole punching with STUN}
-All algorithms here make use of the same principle, UDP hole punching. NAT separates internal
-IP address space from the outer world, thus internet hosts cannot send packet to the 
-internal host directly. Communication has to be initiated from the inside\footnote{Ignoring 
+All algorithms here make use of the same principle, \emph{UDP hole punching}. 
+Internet hosts cannot send packet to the internal host directly. 
+Communication has to be initiated from the inside\footnote{Ignoring 
 pre-set port forwarding, DMZ, uPnP, etc\dots}, what is problematic if two nodes behind 
@@
 
-By default a packet 
-coming from the internet (\textit{IP\textsubscript{A},Port\textsubscript{A}}) to the internal 
-network is dropped on the NAT unless there is a valid mapping in NAT allocation table.
+Assume two parties, A and B with addresses IP\textsuperscript{A} and IP\textsuperscript{B} respectively,
+A and B are behind public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$
+respectively. By default an internal host does not known it's external address and NAT type thus technique like 
+STUN\footnote{Core idea: publicly available service running on 2 IP addresses and 2 ports 
+answering questions like: what is my IP address and port? Change IP address of response if I'll get it, etc...} 
+is used to learn it. In case of incremental NAT it is also needed to determine $\Delta$ and the next port
+that is likely to be allocated for a new connection $\text{PORT}^{\text{A}}_{\text{nat}}$ for A and $\text{PORT}^{\text{B}}_{\text{nat}}$
+for B. The period between determining this information, sending to the other party and starting the algorithm is
+denoted as an \emph{initialization phase} of the NAT traversal algorithm.
+
+UDP hole punching works as follows:
+\begin{compactitem}
+ \item [1.] $A: \; IP^A_{in}:PORT^A_{in} \longrightarrow IP^B_{nat}:PORT^B_{nat}$ \\
+Locally a~new mapping is created ($PORT^A_{nat}$), packet is dropped on B's~NAT.
+ \item [2.] $B: \; IP^A_{nat}:PORT^A_{nat} \longleftarrow  IP^B_{in}:PORT^B_{in}$ \\
+Locally a~new mapping is created ($PORT^B_{nat}$), packet reaches A using allocation created in step~1.
+ \item [3.] $A: \; IP^A_{in}:PORT^A_{in} \longrightarrow IP^B_{nat}:PORT^B_{nat}$ \\
+Packet reaches B using allocation created in in step~2.
+\end{compactitem}
+
+% By default a packet 
+% coming from the internet (\textit{IP\textsubscript{A},Port\textsubscript{A}}) to the internal 
+% network is dropped on the NAT unless there is a valid mapping in NAT allocation table.
 
@@
 The core idea of NAT traversal is to predict the next allocated external port by NAT, thus from this perspective 
-it is important to model how NAT's internal state changes over time, i.e. how many external ports 
-were allocated between two time intervals. Thus the workload is important parameter for predicting allocated port.
+it is important to model how NAT's internal state changes over time, i.e., how many external ports 
+were allocated in a given time interval. Thus the workload is important parameter for predicting allocated port.
 
@@
 \end{mydef}
+
+Moreover we assume incremental NAT with $\Delta=1$ on both sides. If we sample the NAT 
+state\footnote{One new STUN request to determine current port number. Creates new NAT allocation.}
+each $T=10$~milliseconds the NAT state can be modeled as a random process $\{C_i | C_i \in \mathbb{N}, C_i \geq i, i\geq0\}$ where:
+\[
+C_i = \begin{dcases*}
+         0 & if $i=0$ \\
+         C_{i-1} + 1 + X_i, \; X_i \sim Po(\lambda T) & otherwise 
+        \end{dcases*}
+\]
 
@@
 
+ToDo: give an argument that algorithm in each step computes expected value on port number in given time...
+
 \subsection{Algorithm IV - Poisson sampling}
@@
 
+\section{Evaluation}
+ToDo;
+\begin{itemize}
+ \item ToDo: algorithm success rate evaluation w.r.t. $\lambda, T$, graphs, graphs, graphs, \dots
+ \item ToDo: algorithm pros/cons w.r.t. system requirements, applicability
+ \item ToDo: fitting model to the real data from MU network
+ \item ToDo: evaluating algorithms on real data from MU network
+\end{itemize}
+
+\section{Discussion}
+ToDo; What can be done better, different settings, different $\lambda$ on both sides.
+
+\section{Conclusion}
+ToDo; New algorithm
+
 % -----

### 2013-12-07 paper.tex
expected value

@@
 \]
+Rewritten as $C_i = i + \sum_{j=1}^{i}X_i$. Expected value $E[C_i] = i + E[\sum_{j=1}^{i}X_i] = i (1+\lambda T)$.
 
@@
 Expected port value algorithm
-\[
-u \in \mathcal{U}, s \in \mathcal{S}, \lambda \in \mathcal{M}:\;
-\mathcal{A}(u, s, \lambda) = (0, s(1 + \lambda T))
-\]
+\begin{align*}
+& u \in \mathcal{U}, s \in \mathcal{S}, \lambda \in \mathcal{M}:\\
+& \mathcal{A}(u, s, \lambda) = (0, s(1 + \lambda T))
+\end{align*}
 \end{mydef}
 
-ToDo: give an argument that algorithm in each step computes expected value on port number in given time...
+In step $i$ algorithm tries the expected value for $C_i$. Since the $C_i$ has unimodal distribution
+the expected value is also most probable value and the algorithm uses maximum likelihood approach.
 
@@
 
+Success rate of the algorithm strongly relies on accurate measurements of the network 
+properties the client resides in.  Usually
+
 \section{Conclusion}

### 2013-12-10 paper.tex
algI benefits, port pool exhaustion

@@
 
-\section{Related work}
+% \section{Related work}
 Some interesting algorithms for traversing symmetric NAT were proposed by Y. Takeda in \citep{takeda}, but the case both clients are behind symmetric NAT
@@
 Rewritten as $C_i = i + \sum_{j=1}^{i}X_i$. Expected value $E[C_i] = i + E[\sum_{j=1}^{i}X_i] = i (1+\lambda T)$.
+
+\paragraph{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected in this time interval} is assumed to be 3~minutes, 
+what is common value in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~s on average. 
+If it is higher the NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$} pretty fast.
 
@@
 \section{Evaluation}
+
+\paragraph{Algorithm I and II comparison.} 
+Algorithm I has following benefits over II:
+\begin{compactitem}
+ \item works with high success rate for $\lambda \in [0, 0.035] \sim 35$ new connections in 1~s on average.
+ \item can be stopped as soon as connection is established, overhead is proportional to $\lambda$.
+ \item single source port is used, only one listening thread is needed to implement
+it\footnote{NAT is abused to multiplex multiple connections to one source port} compared to the Alg. II.
+This fact makes practical implementation simple and lightweight w.r.t. needed system resources.
+ \item is $\lambda$-invariant. Provided network load is low no additional measurements are needed.
+\end{compactitem}
+
 ToDo;

### 2013-12-11 paper.tex
evaluation, discussion, conslusion

@@
 \usepackage{fixltx2e}
+% \usepackage{wrapfig}
+\usepackage{tikz}
+\usepackage{float}
+
 \newtheorem{mydef}{Definition}
@@
 for B. The period between determining this information, sending to the other party and starting the algorithm is
-denoted as an \emph{initialization phase} of the NAT traversal algorithm.
+denoted as an \emph{initialization phase} or a \emph{silent period}\footnote{Since the algorithm itself does nothing} of the NAT traversal algorithm.
 
-UDP hole punching works as follows:
+UDP hole punching works as follows:\\
 \begin{compactitem}
@@
 parameters and additional information obtained in initial phase). 
-In our setting $\mathcal{M} = \{ (\lambda) | \lambda > 0 \}$. For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
+In our setting $\mathcal{M} = \{ (\lambda, t_s) | \lambda~>~0, t\geq~0 \}$. Value $t_s$ is length of the silent period. %For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
 \item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
@@
 \end{mydef}
-If third parameter is not important for the algorithm it can be ignored in notation.
+If the third parameter is not important for the algorithm it can be ignored in the notation.
+
+Prior algorithm run it is needed to 
+determine NAT state (UDP hole punching), this is assumed to be done by some technique like STUN. For simplicity we assume the port numbers
+starts from $0$. The length of the silent period is also known to the algorithm prior execution in order to estimate the current state of 
+the NAT, e.g., difference between the last measurement and just before algorithm execution.
 
@@
 \begin{align*}
-& u \in \mathcal{U}, s \in \mathcal{S}, \lambda \in \mathcal{M}:\\
-& \mathcal{A}(u, s, \lambda) = (0, s(1 + \lambda T))
+& u \in \mathcal{U}, s \in \mathcal{S}, (\lambda, t_s) \in \mathcal{M}:\\
+& \mathcal{A}(u, s, (\lambda, t_s)) = (0, \lambda t_s + s(1 + \lambda T))
 \end{align*}
@@
 \begin{align*}
-& u \in \mathcal{U}, s \in \mathcal{S}, \lambda \in \mathcal{M}:\\
-& \mathcal{A}(u, s, \lambda) = (0, \text{Poisson}(\lambda T (1+s C(\lambda)))
-\end{align*} where $\text{Poisson}(\mu) = X \sim Po(\mu)$ samples Poisson distribution with
-parameter $\mu$, function $C(\lambda)$ gives a constant 
-coefficient\footnote{This function was found empirically by maximizing success rate}.
+& u \in \mathcal{U}, s \in \mathcal{S}, (\lambda, t_s) \in \mathcal{M}:\\
+& \mathcal{A}(u, s, (\lambda, t_s)) = (0, \lambda t_s + \text{P}(\lambda T (1+s C(\lambda T)))
+\end{align*} where $\text{P}(\mu) = X \sim Po(\mu)$, samples Poisson distribution with
+parameter $\mu$, coefficient function $C(\lambda T)$ gives a constant 
+coefficient\footnote{This function was found empirically by maximizing the success rate}.
+Approximations of the $C(\lambda T)$ by an inverse logarithm and a quartic model:
+\begin{align*}
+C(x) = & ({0.163321 \cdot  ln(64.2568 \cdot x)})^{-1}\\
+C(x) = & (0.172876+1.28162\cdot x-1.41256\cdot x^2 \\
+       & +0.825093\cdot x^3-0.184726\cdot x^4)^{-1}
+\end{align*}
 \end{mydef}
+% 
+\subsection{Silent period in algorithms.} If the length of the silent period is too big
+the algorithm can have problems with coping with a gap that occurred in NAT ports. Thus 
+Algorithms III and IV estimates the starting offset caused by the silent period as an
+expected value from $E[Po(\lambda \cdot t_1)] = \lambda\cdot t_1$.
 
@@
 \begin{compactitem}
- \item works with high success rate for $\lambda \in [0, 0.035] \sim 35$ new connections in 1~s on average.
- \item can be stopped as soon as connection is established, overhead is proportional to $\lambda$.
- \item single source port is used, only one listening thread is needed to implement
+ \item Works with high success rate for $\lambda \in [0, 0.035] \sim 35$ new connections in 1~s on average.
+ \item Can be stopped as soon as connection is established, overhead is proportional to~$\lambda$.
+ \item Single source port is used, only one listening thread is needed to implement
 it\footnote{NAT is abused to multiplex multiple connections to one source port} compared to the Alg. II.
-This fact makes practical implementation simple and lightweight w.r.t. needed system resources.
- \item is $\lambda$-invariant. Provided network load is low no additional measurements are needed.
+Due to this fact a practical implementation is simple and lightweight w.r.t. needed system resources.
+ \item Is $\lambda$-invariant. Provided network load is low no additional measurements are needed.
 \end{compactitem}
 
-ToDo;
+\paragraph{Algorithms simulation.}
+
+\begin{figure}[H]
+% \begin{wrapfigure}{r}{0.5\textwidth}
+% \begin{figure*}
+% \begin{center}
+% \leavevmode
+% \centerline
+%{\scalebox{0.40}{\includegraphics[width=\textwidth]{alg_1.pdf}}}
+{\scalebox{0.43}{\includegraphics[trim=50 0 0 0]{alg_1.pdf}}}
+% \end{center}
+\caption{Algorithm evaluation in a simulation, $T=10~ms$. Number of a simulation rounds for one parameter setting is 1000.
+ A, B, C correspond to the Algorithm I, II and III respectively.
+ D, E correspond to the Algorithm IV with optimized $C(\lambda T)$ by an exhaustive search in D case, 
+ in E case $C(\lambda T)$ is approximated by a polynomial.}
+\label{fig:alg_eval}
+\end{figure}
+% \end{figure*}
+% \end{wrapfigure}
+
+\begin{figure}[H]
+{\scalebox{0.43}{\includegraphics[trim=40 0 0 0]{alg_steps.pdf}}}
+\caption{Steps of the algorithms needed to success on average w.r.t. $\lambda$. Algorithms denoted in a same way as in figure \ref{fig:alg_eval}}
+\label{fig:alg_steps}
+\end{figure}
+
+In figure \ref{fig:alg_eval} is depicted result of a simulation of 
+the algorithms, for different $\lambda$. Figure \ref{fig:alg_steps} depicts number of steps needed 
+to succeed on average. From figures it is visible Alg. I is more suitable for networks with low 
+workload since has $100\%$ success rate for $\lambda \in [0, 0.035]$. Outside this interval 
+Alg. I is not working and another solution has to be applied. The problem is new connections are
+created faster than algorithm progresses.
+
+Alg. II is set to have fixed $\Delta=900$, in order to maximize the success rate. The disadvantage of this
+algorithm is that it waits on a particular destination port and if it is taken by another connection
+the algorithm fails.
+
+Alg. IV has high success rate for 
+noisy networks but is very sensitive on a $\lambda$ estimation what decreases practical usability 
+if $\lambda$ changes quickly over time (non-homogeneous Poisson process). Closed form formula for $C(\lambda T)$ 
+optimizing success rate and analytic derivation of the algorithm is still an open question.
+
 \begin{itemize}
- \item ToDo: algorithm success rate evaluation w.r.t. $\lambda, T$, graphs, graphs, graphs, \dots
- \item ToDo: algorithm pros/cons w.r.t. system requirements, applicability
  \item ToDo: fitting model to the real data from MU network
@@
 \section{Discussion}
-ToDo; What can be done better, different settings, different $\lambda$ on both sides.
+For simplicity here we assumed the $\lambda$ is the same on both sides, the new challenges 
+can arise if we allow different parameters on both sides that we haven't studied yet.
+
+Another open question is to find a model that optimizes given task, i.e., establishing a direct
+connection. We proposed a start step, to model the same process that occurs on the network, i.e.,
+Poisson process, but this method requires setting a coefficient function $C(\lambda T)$ for which 
+we don't have exact analytic explanation. This method is also very sensitive to parameter fluctuations
+and thus not very practical for a real world networks.
+
+In our setting we assume $T=10~ms$ in order to minimize the speed of the evolution of a Poisson process.
+Such short sending interval can cause some packets may get dropped on the path to the destination. But the 
+important is whether they reach the local NAT in order to create a mapping in allocation table. As proposed 
+in \citep{Wang:2006:RSN:1156422.1156550} it is possible to run the algorithm with $T=10~ms$ to create a mapping
+and then re-run the algorithm with higher $T$, e.g., $T=100~ms$ in order to deliver packets to the destination
+using created mapping and holes created in the previous round.
 
 Success rate of the algorithm strongly relies on accurate measurements of the network 
-properties the client resides in.  Usually
+properties the client resides in. They can be measured prior connection establishment 
+from the point the client connects to the network, as is usually done in ICE and gather 
+more information, build better network model and so on in this way.
+
+Another problematic thing is if this algorithm is followed by a multiple host pairs on
+the same network in the same time. They would artificially increase the network workload
+causing it wouldn't follow Poisson distribution rendering our algorithms as not usable in
+this case.
 
 \section{Conclusion}
-ToDo; New algorithm
+The aim of this work was to study problem of the NAT traversal from statistical point of view,
+and to provide an algorithm with high success rate for networks with low workload. We proposed
+several algorithms solving this problem but there are still some open questions remaining.
 

### 2013-12-15 paper.tex
vashek comments, algorithm successful run

@@
 % BibTex for bibliography
-\usepackage[numbers]{natbib}
+\usepackage[square,numbers]{natbib}
 %\usepackage[square,sort&compress,authoryear]{natbib}
@@
 \usepackage{float}
+\usepackage{multirow}
+
+\usepackage{algorithm}% http://ctan.org/pkg/algorithm
+\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
+\usepackage[compatibility=false]{caption}% http://ctan.org/pkg/caption
 
@@
 	\fontsize{24pt}{10pt}\selectfont
-	\textbf{Traversing Symmetric NAT with predictable port allocation function}
+	\textbf{Traversing symmetric NAT with predictable port allocation function}
 	}	
@@
 	\normalsize	Masaryk University, Brno \\
-	\normalsize	\href{mailto:xklinec@mail.muni.cz}{xklinec@mail.muni.cz}
+	\normalsize	\href{mailto:xklinec@fi.muni.cz}{xklinec@fi.muni.cz}
 	\and
@@
 	\normalsize	Masaryk University, Brno \\
-	\normalsize	\href{mailto:matyas@mail.muni.cz}{matyas@mail.muni.cz}
+	\normalsize	\href{mailto:matyas@fi.muni.cz}{matyas@fi.muni.cz}
 	\vspace{-5mm}
@@
 it's parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
-between communicating parties is necessary. If P2P connection is not possible to be established relay 
+between communicating parties is necessary. If P2P connection is not possible to be established, relay 
 servers has to be used (TURN~\citep{rfc5766} protocol is used). The presence 
@@
 
+\section{Existing techniques}
+Many different solutions for traversing non-symmetric NATs exist e.g., NAT traversal in Internet Key Exchange~\citep{rfc3947} or 
+Interactive Connectivity Establishment~(ICE)~\citep{rfc5245}. %Terredo\ignore{tunneling}~\citep{rfc4380}.
+
+Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}. But in general 
+it is not widely deployed and it cannot be relied upon in a general case.
+
+Most of the applications use ICE\ignore{~\citep{rfc5245}} nowadays, what is a very powerful toolbox for traversing 
+various types of NAT, however does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for it's operation.
+
+% \section{Related work}
+Some interesting algorithms for traversing symmetric NAT were proposed by Y. Takeda in \citep{takeda}, but the case both clients are behind symmetric NAT
+with port sensitive allocation function is not covered. It mentions that sending more packets does not increase a probability of establishing a new connection.
+Our motivation is to show that it is actually possible.
+
+RFC 5128 \citep{rfc5128} maps current state of the art in NAT traversal, but does not cover our case. 
+The research by Wei~\emph{et.~al.} in \citep{wei}., is devoted to symmetric NAT traversal but their algorithm is using another technique compared
+to our approach. The downside of their algorithm is it requires Time To Live~(TTL) modification of the IP~packet. This is usually considered
+as a privileged operation in the operating system and super-user privileges are required for this modification.
+
+In our work we study the situation where two parties, A,B tries to establish a direct connection, both of them are behind 
+a symmetric NAT (possibly cascade of such NATs) with port sensitive, predictable allocation function. Another scenarios with
+predictable allocation function are quite easy to solve and covered in previous research.
+
 \section{Network Address Translation}
-The notation we will be using for the rest of the paper follows.
+The notation used for the rest of the paper follows.
 
@@
 % Rule is created once the connection from the internal host is initiated to an arbitrary external address.
-Filtering rule is
+Filtering rule is:
 \begin{align*}
@@
 
-\emph{Allocation function (AF)} says how is the port allocated for new connection chosen. Assume $\mathbb{AT}=\emptyset$, for simplicity. 
-AF can be \emph{predictable}, e.g. \emph{incremental}. If previously allocated port was $\text{PORT\textsubscript{nat}}$
+\emph{Allocation function (AF)} defines port allocation rules for new connections (not in $\mathbb{AT}$). Assume $\mathbb{AT}=\emptyset$, for simplicity. 
+AF can be either \emph{predictable} or \emph{random}. In our setting predictable AF is assumed, e.g., \emph{incremental}. 
+If previously allocated port was $\text{PORT\textsubscript{nat}}$
 the next port will be $\text{PORT\textsubscript{nat}}+\Delta$, where $\Delta=1$ usually. Another type of AF is \emph{random}.
@@
 \emph{Port sensitive allocation function} creates new allocation if a destination address and a port differs from the existing
-allocations from the same source\footnote{for each new connection is allocated new a NAT port.}:
+allocations from the same source:
 \begin{align*}
@@
 
-\section{Existing techniques}
-Many different solutions for traversing non-symmetric NATs exist e.g., NAT traversal in Internet Key Exchange~\citep{rfc3947} or Terredo\ignore{tunneling}~\citep{rfc4380}.
-
-Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}. But in general 
-it is not widely deployed and it cannot be relied upon.
-
-Most of the applications use Interactive Connectivity Establishment~(ICE)~\citep{rfc5245} nowadays, what is a very powerful toolbox for traversing 
-various types of NAT, however does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN~\citep{rfc5766} for it's operation.
-
-% \section{Related work}
-Some interesting algorithms for traversing symmetric NAT were proposed by Y. Takeda in \citep{takeda}, but the case both clients are behind symmetric NAT
-with port sensitive allocation function is not covered. It mentions that sending more packets won't increase probability of establishing the new connection.
-Our motivation is to show that it is actually possible.
-
-The \citep{rfc5128} maps current state of the art in NAT traversal. But does not cover our case.
-
-Also the academic research was done in this field, for example by Wei~\emph{et.~al.} in \citep{wei}., but their algorithm is using another technique compared
-to our approach. The downside of their algorithm is it requires Time To Live~(TTL) modification of the IP~packet. This is in majority of cases considered
-as a privileged operation in operating system and sufficient privileges are required for this modification.
-
-In our work we study the situation where two parties, A,B tries to establish a direct connection, both of them are behind 
-a symmetric NAT (possibly cascade of such NATs) with port sensitive, predictable allocation function. Another scenarios with
-predictable allocation function are quite easy to solve and covered in previous research.
-
 % \section{Related work II}
@@
 Internet hosts cannot send packet to the internal host directly. 
-Communication has to be initiated from the inside\footnote{Ignoring 
-pre-set port forwarding, DMZ, uPnP, etc\dots}, what is problematic if two nodes behind 
+Communication has to be initiated from the inside\footnote{Not taking
+pre-set port forwarding, DMZ, uPnP, etc... into consideration. }, what is problematic if two nodes behind 
 NAT wants to communicate.
@@
 for B. The period between determining this information, sending to the other party and starting the algorithm is
-denoted as an \emph{initialization phase} or a \emph{silent period}\footnote{Since the algorithm itself does nothing} of the NAT traversal algorithm.
+denoted as an \emph{initialization phase} or a \emph{silent period}\ignore{\footnote{Since the algorithm itself does nothing.}} of the NAT traversal algorithm.
 
@@
 Moreover we assume incremental NAT with $\Delta=1$ on both sides. If we sample the NAT 
-state\footnote{One new STUN request to determine current port number. Creates new NAT allocation.}
+state\footnote{One new STUN request to determine current port number, it creates a new NAT allocation record.}
 each $T=10$~milliseconds the NAT state can be modeled as a random process $\{C_i | C_i \in \mathbb{N}, C_i \geq i, i\geq0\}$ where:
@@
 \]
-Rewritten as $C_i = i + \sum_{j=1}^{i}X_i$. Expected value $E[C_i] = i + E[\sum_{j=1}^{i}X_i] = i (1+\lambda T)$.
+Rewritten as $C_i = i + \sum_{j=1}^{i}X_j$. Expected value $E[C_i] = i + E[\sum_{j=1}^{i}X_i] = i (1+\lambda T)$.
 
-\paragraph{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected in this time interval} is assumed to be 3~minutes, 
+\paragraph{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected in this time interval.} is assumed to be 3~minutes, 
 what is common value in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~s on average. 
-If it is higher the NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$} pretty fast.
+If it is higher the NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$.} pretty fast.
 
@@
 \begin{compactitem}
-\item $\mathcal{U}=\{0,1\}$ is set of communicating parties
-\item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$ is a step of the algorithm
+\item $\mathcal{U}=\{0,1\}$ is set of communicating parties.
+\item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$ is a step of the algorithm.
 \item $\mathcal{M}$ is set of vectors of parameters of the environment (i.e., network model 
@@
 \item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
-are sets of source and destination ports respectively
+are sets of source and destination ports respectively.
 \end{compactitem}
@@
 \paragraph{Intuition.} 
-If we have $A(0,1) = (2,3)$ it means party $0$ in step $1$ sent packet from source port $2$ to destination port $3$
+Line $A(0,1) = (2,3)$ means party $0$ in step $1$ sent packet from source port $2$ to destination port $3$
 of the party $1$.
@@
 
-\paragraph{Party change.} 
-We define unary operation on communicating parties. 
-\[
-a \in \mathcal{U},  \neg a = \begin{dcases*}
-         0 & when $a=1$\\
-         1 & when $a=0$
-        \end{dcases*}
-\]
+%\paragraph{Party change.} 
+%We define unary operation on communicating parties. 
+%\[
+%a \in \mathcal{U},  \neg a = \begin{dcases*}
+%         0 & when $a=1$\\
+%         1 & when $a=0$
+%        \end{dcases*}
+%\]
 
-\subsection{Algorithm I - baby-step, giant-step}
+\paragraph{Algorithm log \& successful run.} IP addresses of the hosts and NAT are known and fixed prior algorithm run thus omitted
+from the notation for simplicity. For the following text is assumed maximum number of steps of the algorithm $s_{max}=1000$.
+
+\begin{mydef}
+\label{def:log}
+Algorithm log $\mathcal{L}^A_{\mathcal{A}}$ is a set of tuples $(p_1, p_2, p_3) \in \mathcal{L}^A_{\mathcal{A}}$ produced
+by NAT traversal algorithm $\mathcal{A}$ execution on host $A$, where: \\
+%$\mathcal{A}(0, s, (\lambda, t_s)) = (p_1, p_2)$ 
+%are generated by $\mathcal{A}$ for some parameters $\lambda, t_s$ in some step $s$. The $p_3$ is 
+%an allocated port by $NAT^A$ for request $(p_1, p_2)$ in a given step. Algorithm log for host $B$ is defined analogically. 
+%Furthermore from our setting the 
+%following formula holds:\footnote{From definition of a port sensitive symmetric NAT.} %$\mathcal{L^{A}} \subseteq PORT_{in}^{A} \times PORT_{nat}^{B} \times PORT_{nat}^{A}$,
+
+\begin{compactitem}
+ \item $(p_1, p_2) = \mathcal{A}(0, s, (\lambda, t_s)) $ are generated by the algorithm $\mathcal{A}$ for some 
+       parameters $(\lambda, t_s)$ in some step $s$ of the algorithm $\mathcal{A}$.
+ \item $p_3$ is an allocated port by $NAT^A$ for request $(p_1, p_2)$ in a given step $s$. In our network model
+       $p_3=C_s=C_{s-1} + 1 + X, X~\sim~Po(\lambda T)$.
+ \item Algorithm log for host $B$ is defined analogically.
+ \item Furthermore from our setting the following formula holds:\footnote{From definition of a port sensitive symmetric NAT.}
+\end{compactitem} %
+%
+% \begin{compactitem}
+% \item $p_1$ is a source port.
+% \item $p_2$ is a destination port.
+% \item $p_3$ is an allocated port by local NAT for this request.
+% \item following formula holds\footnote{from definition of a port sensitive symmetric NAT}
+% \end{compactitem}
+%
+\begin{align*}
+& (p_1, p_2, p_3) \in \mathcal{L}^A_{\mathcal{A}}: \forall (p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{A}_{\mathcal{A}} :\\
+& p_3 = p_3^{\prime} \Leftrightarrow \left( p_1=p^{\prime}_1 \wedge p_2=p^{\prime}_2 \right)
+\end{align*}
+\end{mydef}
+
+\begin{mydef}
+\label{def:match}
+Algorithm $\mathcal{A}$ run is successful, i.e., established a direct connection between $A,B$ if\footnote{Derived from UDP hole punching.}:
+\begin{align*}
+& MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}}) \Leftrightarrow \exists(p_1, p_2, p_3) \in \mathcal{L}^{A}_{\mathcal{A}}: \\
+& \exists(p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{B}_{\mathcal{A}}: p_2 = p_3^{\prime} \wedge p_2^{\prime} = p_3
+%
+%  \in \mathcal{L}^A_{\mathcal{A}}: \forall (p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{A}_{\mathcal{A}} :\\
+%& p_3 = p_3^{\prime} \left( \Leftrightarrow p_1=p^{\prime}_1 \wedge p_2=p^{\prime}_2 \right)
+\end{align*}
+\end{mydef}
+The task is to find an algorithm $\mathcal{A}$ such that the probability the $MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}})$
+holds is maximized for minimal number of steps needed.
+
+\subsection{Algorithm I -- baby-step, giant-step}
 We propose the algorithm with high success rate for small network loads, i.e.,
@@
 
-\subsection{Algorithm II - fixed destination port}
+\subsection{Algorithm II -- fixed destination port}
 Algorithm with fixed destination port described in by \citep{Wang:2006:RSN:1156422.1156550}
@@
 
-\subsection{Algorithm III - expected port value}
+\subsection{Algorithm III -- expected port value}
 Some more effective algorithms can be designed if the random process the network follows is known (and their
@@
 
-\subsection{Algorithm IV - Poisson sampling}
+\subsection{Algorithm IV -- Poisson sampling}
 Algorithm with the best success rate in our simulation is based on sampling Poisson distribution.
 The main idea is to let the algorithm simulate the Poisson process that the network of the
-other party follows.
+other party follows. Duplicates are removed since it does not lead to new port allocation.
 
@@
 Poisson sampling algorithm
-\begin{align*}
-& u \in \mathcal{U}, s \in \mathcal{S}, (\lambda, t_s) \in \mathcal{M}:\\
-& \mathcal{A}(u, s, (\lambda, t_s)) = (0, \lambda t_s + \text{P}(\lambda T (1+s C(\lambda T)))
-\end{align*} where $\text{P}(\mu) = X \sim Po(\mu)$, samples Poisson distribution with
+
+% \captionof{algorithm}{Euclid’s algorithm}\label{algo1}
+
+\begin{algorithmic}[1]
+  \algtext*{EndWhile}% Remove "end while" text
+  \algtext*{EndIf}% Remove "end if" text
+  \algtext*{EndFunction}% Remove "end if" text
+  \newcommand{\LineIf}[2]{ \State \algorithmicif\ {#1}\ \algorithmicthen\ {#2}}% \algorithmicend\ \algorithmicif }
+  \Function{init4}{$\lambda, T$} %\Comment{The g.c.d. of a and b}
+    \State $B\gets [], s \gets 0$
+    %\State $s\gets 0$
+    \While{$len(B)< s_{max}$}
+      \State $x\gets \text{P}(\lambda T (1+s C(\lambda T))$
+      \State $s\gets s+1$
+      \LineIf{$x \notin B$}{$B.append(x)$} %\If {$x \notin B$} \State $B.append(x)$ \EndIf
+    \EndWhile
+    \State \textbf{return} $B$
+    %\Return $B$
+   \EndFunction\\
+
+   \State $M\gets init4(\lambda,T)$\\
+   \State $u \in \mathcal{U}, s \in \mathcal{S}, (\lambda, t_s) \in \mathcal{M}:$
+%     \State $\mathcal{A}(u, s, (\lambda, t_s)) = (0, \lambda t_s + M[s])$
+   \Function{A}{$u, s, (\lambda, t_s)$} %\Comment{The g.c.d. of a and b}
+    %\LineIf{$s = 0$}{$M = init4(\lambda,T)$} %\If {$x \notin B$} \State $B.append(x)$ \EndIf
+    \State \textbf{return} $(0, \lambda t_s + M[s])$
+   \EndFunction
+\end{algorithmic}
+%\begin{align*}
+%& u \in \mathcal{U}, s \in \mathcal{S}, (\lambda, t_s) \in \mathcal{M}:\\
+%& \mathcal{A}(u, s, (\lambda, t_s)) = (0, \lambda t_s + \text{P}(\lambda T (1+s C(\lambda T)))
+%\end{align*} 
+where $\text{P}(\mu) = X \sim Po(\mu)$, samples Poisson distribution with
 parameter $\mu$, coefficient function $C(\lambda T)$ gives a constant 
-coefficient\footnote{This function was found empirically by maximizing the success rate}.
+coefficient\footnote{This function was found empirically by maximizing the success rate in simulations.}.
 Approximations of the $C(\lambda T)$ by an inverse logarithm and a quartic model:
@@
 Algorithms III and IV estimates the starting offset caused by the silent period as an
-expected value from $E[Po(\lambda \cdot t_1)] = \lambda\cdot t_1$.
+expected value $E[Po(\lambda \cdot t_1)] = \lambda\cdot t_1$.
 
@@
 \begin{compactitem}
- \item Works with high success rate for $\lambda \in [0, 0.035] \sim 35$ new connections in 1~s on average.
+ \item Works with high success rate for $\lambda \in [0, 0.035] \sim 35$ new connections in 1~sec on average.
  \item Can be stopped as soon as connection is established, overhead is proportional to~$\lambda$.
  \item Single source port is used, only one listening thread is needed to implement
-it\footnote{NAT is abused to multiplex multiple connections to one source port} compared to the Alg. II.
+it\footnote{NAT is abused to multiplex multiple connections to one source port.} compared to the Alg. II.
 Due to this fact a practical implementation is simple and lightweight w.r.t. needed system resources.
@@
 %{\scalebox{0.40}{\includegraphics[width=\textwidth]{alg_1.pdf}}}
-{\scalebox{0.43}{\includegraphics[trim=50 0 0 0]{alg_1.pdf}}}
+{\scalebox{0.43}{\includegraphics[trim=40 0 0 0]{alg_1.pdf}}}
 % \end{center}
@@
 {\scalebox{0.43}{\includegraphics[trim=40 0 0 0]{alg_steps.pdf}}}
-\caption{Steps of the algorithms needed to success on average w.r.t. $\lambda$. Algorithms denoted in a same way as in figure \ref{fig:alg_eval}}
+\caption{Steps of the algorithms needed to success on average w.r.t. $\lambda$. Algorithms denoted in a same way as in figure \ref{fig:alg_eval}.}
 \label{fig:alg_steps}
@@
 
+\section{Real data modeling}
+Data collected on university network by NetFlow probes were studied to test the correspondence to the Poisson process
+w.r.t. newly created connections/allocations on NAT. NAT was simulated in our program using
+NetFlow data to provide network traffic to the NAT. NetFlow data capture 1 hour of a network communication in two different
+day hours (03 -- low peak, 13 -- high peak).
+
+NAT state (newly opened connections in window $T=10, 50, 100~ms$)
+was sampled multiple times in a row, Poisson and Negative Binomial distributions were fitted using Maximum Likelihood Estimation (MLE)
+and goodness-of-fit test with Pearson $\chi^2$ test was performed with confidence level $\alpha=0.05$.
+
+\begin{table*}[t]
+  \centering
+  % set, T, sample size, PoFit, NB Fit, sample mean, variance of sample mean, sample variance
+  \begin{tabular}{| l | r | r | r | r | r | r | r |} % 
+    \hline
+    Set & T [ms] & Samples & Poisson r. [\%] & NBinom r. [\%] & $\overline{m}_{E[X]}$ &  $V[\overline{m}_{E[X]}]$  & $\overline{m}_{V[X]}$ \\ \hline \hline 
+    \multirow{7}{*}{03} & \multirow{3}{*}{10}   & 100		& 2.977		& 1.667		& 0.5121        & 0.0228                    & 0.5321 \\ \cline{3-8}
+      &      & 500      	& 5.637		& 0.920		& 0.4851        & 0.0086        & 0.5146 \\  \cline{3-8}
+      &      & 1000     	& 10.294	& 1.579		& 0.4831        & 0.0068        & 0.5142 \\  \cline{2-8}
+      & \multirow{3}{*}{50}     & 100     	& 6.543		& 2.276 	& 2.3707	& 0.2373 & 2.9969 \\ \cline{3-8}
+      &      & 500     		& 25.180	& 7.194		& 2.3871        & 0.1053        & 3.1179 \\ \cline{3-8}
+      &      & 1000    		& 52.174	& 23.188	& 2.3920        & 0.0696        & 3.1576 \\ \cline{2-8}
+      & \multirow{2}{*}{100}  	& 100		& 8.914		& 3.911		& 4.7263        & 0.7031                    & 7.1496 \\ \cline{3-8}
+      &      & 500     		& 57.746	& 9.859		& 4.7668	& 0.2876 	& 7.4667 					\\ \hline
+
+    \multirow{2}{*}{13}  & \multirow{2}{*}{10}     & 100     & 5.191      & 1.824     & 2.6438       &  0.2695    & 3.0020 \\ \cline{3-8}
+      &      & 500     & 20.368     & 2.829     & 2.6579       &  0.1025                    & 3.1495 \\ \hline
+  \end{tabular}
+  \caption{Results from NAT simulation on the real data captured from the university network. 
+    The meaning of columns from the left is following: data set name, sampling time, number of samples collected,
+    \% of cases rejecting hypothesis about Poisson distribution,  
+    \% of cases rejecting hypothesis about Negative Binomial distribution,
+    sample mean of sample mean of port samples,
+    sample variance of sample mean of port samples,
+    sample mean of variance of port samples.}
+  \label{tab:1}
+\end{table*}
+
+Regarding data from set 13, during activity peak during the day, we conclude that if a number of samples 
+is lower, e.g., 100, the process of a creating a new connection on NAT can 
+be modeled as a time homogeneous Poisson process successfully (hypothesis rejected in $5.191\%$ of cases).
+The time window the network was monitored in this case is $T \cdot 100 = 1000~ms$.
+
+If the number of samples is higher, e.g., 500, the time window of studied network traffic is longer and
+in a real world scenario the probability that $\lambda$ is constant in time is low with increasing the time
+window. Due to this time homogeneous model is not strong enough, data is 
+overdispersed what causes failure to fit Poisson distribution (hypothesis rejected in $20.368\%$ of cases). As suggested
+in literature, Negative Binomial model fits better for overdispersed\footnote{The data overdispersion may be caused by dependent events on the network e.g., time-triggered updates downloading, etc...}
+data (hypothesis rejected in $2.829\%$ of cases). 
+The open question is how to make use of this model in NAT traversal.
+
+For 100 samples the match with the model is satisfactory. Provided algorithm is able to run completely (including learning $\lambda$),
+the proposed algorithms should work like in a simulation.
+
+Regarding data from set 03, we can conclude that lower network workload causes more stable mean of sampled ports (MLE for Poisson distribution),
+thus the Poisson distribution hypothesis is not rejected even in longer time intervals for an estimated $\lambda$.
+
 \section{Discussion}
@@
 
+%We ignored the cases where there are already some allocations on NAT from older connections
+%on the port interval the NAT traversal algorithm is using.
+
 Another problematic thing is if this algorithm is followed by a multiple host pairs on
@@
 The aim of this work was to study problem of the NAT traversal from statistical point of view,
-and to provide an algorithm with high success rate for networks with low workload. We proposed
-several algorithms solving this problem but there are still some open questions remaining.
+and to provide an algorithm with high success rate for networks with low workload that is easy to 
+implement and requires little system resource. 
+Several algorithms solving this problem but there are still some open questions remaining.
+
+We modeled the network workload as a Poisson process what we verified on real data by hypothesis testing.
+
+We developed an application written in Python that simulates NAT and mentioned algorithms, computes
+statistics for port distributions on NAT and simulates algorithms for different parameters. It
+may be helpful for another researchers interested in this area so we published it in public domain
+hosted on github [URL:WHERE].
 
@@
 % Bibliography
-\bibliographystyle{unsrtnat}
+%\bibliographystyle{unsrtnat}
+\bibliographystyle{plainnat}
 \bibliography{paper}

### 2013-12-15 paper.tex
some fixes

@@
 	\fancyfoot{}
-	\fancyhead[C]{Security Protocol Workshop $\bullet$ March 2014 $\bullet$ Vol 1}
+	\fancyhead[C]{SPW 2014 $\bullet$ Twenty-second International Workshop on Security Protocols}
 	\fancyfoot[RO,LE]{\thepage}
+% CFP: http://www.wikicfp.com/cfp/servlet/event.showcfp?eventid=33620&copyownerid=44118
 
@@
 \begin{align*}
-& PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
+& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
 &  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
@@
 Algorithm log $\mathcal{L}^A_{\mathcal{A}}$ is a set of tuples $(p_1, p_2, p_3) \in \mathcal{L}^A_{\mathcal{A}}$ produced
-by NAT traversal algorithm $\mathcal{A}$ execution on host $A$, where: \\
+by the NAT traversal algorithm $\mathcal{A}$ execution on the host $A$, where: \\
 %$\mathcal{A}(0, s, (\lambda, t_s)) = (p_1, p_2)$ 
@@
 \mathcal{A}(u, s) = \begin{dcases*}
-         (0, s)  & when $u=0$\\
-         (0, 2s) & when $u=1$
+         (0, s)  & \text{if} $u=0$\\
+         (0, 2s) & \text{if} $u=1$
         \end{dcases*}
@@
 
-Intuitively the source port is kept constant, one side of the protocol scans ports sequentially, another
-side of the protocol scans ports with step 2.
-
 \subsection{Algorithm III -- expected port value}
@@
 The main idea is to let the algorithm simulate the Poisson process that the network of the
-other party follows. Duplicates are removed since it does not lead to new port allocation.
+other party follows. Duplicates are removed since it does not lead to a new port allocation.
 

### 2013-12-16 paper.tex
colaborating with the enemy, introduction, abstract

@@
 \usepackage{multirow}
+%\usepackage{appendices}
 
@@
 \begin{abstract}
-\noindent ToDo; Will be written as the last part of the paper...
+\noindent Network Address Translators cause troubles for VoIP and other P2P services since central servers are needed for communication.
+A presence of such potentially malicious hosts in a communication path is not desired, mainly due to increased cost, poor link quality 
+and some security consequences.
+Several solution exist for traversing NAT but symmetric case is still problematic. We propose algorithms using single source port for 
+symmetric NAT traversal, each with different properties and applicability. 
+Paper gives a theoretic point of view on the NAT traversal problem by stating problem formally and modeling a network connected to 
+NAT as a random process, in order to evaluate algorithms in a different settings.
+% 
+% ToDo; Will be written as the last part of the paper...
 Lorem ipsum dolor sit amet, consectetur adipiscing elit.
@@
 Network Address Translators (abbreviated as NAT) has been increasing in popularity in recent years 
-mainly due to IPv4 address space depletion, in order to slow it down. On one side NAT offers
-the ability to share a public, globally routable IP address among many internal network hosts but on 
-the other side it often breaks basic end-to-end principle of the Internet and thus favors server-client 
+mainly as a mean of slowing down IPv4 address space depletion. On one hand NAT offers
+the ability to share a public, globally routable IP address among many internal network hosts. On 
+the other hand it breaks basic end-to-end principle of the Internet and thus favors server-client 
 communication model for clients connected behind NAT.
 
-In many applications is peer-to-peer (P2P) connection of crucial importance and the quality of a service depends on 
-it's parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
+In many applications is a peer-to-peer (P2P) connection of crucial importance and the quality of a service depends on 
+its parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
 between communicating parties is necessary. If P2P connection is not possible to be established, relay 
@@
 (increased latency, jitter, lower bandwidth), relay server is a potential point of failure, 
-has to scale with increasing demands of the network, interesting point for attack, etc.
+has to scale with increasing demands of the network, etc...
 Thus relay servers require non-negligible resources what makes them expensive for VoIP providers and
 establishing a direct connection between parties is of high importance.
+
+From an attacker perspective, relay servers are interesting point where to attack thus from security
+standpoint they could be seen as potential enemies trying to sniff and log all 
+communication. NAT (especially symmetric one, used as a carrier-grade border firewall in large networks)
+forces communicating parties to use such central servers
+for communication making monitoring easier for an attacker. Traversal algorithms has to collaborate 
+with central servers in such a way they create a direct link, not using servers any further. 
+% 
+% some way. They use it to create a direct link, not using servers any further.
 
@@
 Most of the applications use ICE\ignore{~\citep{rfc5245}} nowadays, what is a very powerful toolbox for traversing 
-various types of NAT, however does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for it's operation.
+various types of NAT, however does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for its operation.
 
 % \section{Related work}
-Some interesting algorithms for traversing symmetric NAT were proposed by Y. Takeda in \citep{takeda}, but the case both clients are behind symmetric NAT
+Some interesting algorithms for traversing symmetric NAT were proposed by Y.~Takeda in \citep{takeda}, but the case both clients are behind symmetric NAT
 with port sensitive allocation function is not covered. It mentions that sending more packets does not increase a probability of establishing a new connection.
@@
 
-RFC 5128 \citep{rfc5128} maps current state of the art in NAT traversal, but does not cover our case. 
+RFC~5128 \citep{rfc5128} maps current state of the art in NAT traversal, but does not cover our case. 
 The research by Wei~\emph{et.~al.} in \citep{wei}., is devoted to symmetric NAT traversal but their algorithm is using another technique compared
@@
 as a privileged operation in the operating system and super-user privileges are required for this modification.
+
+Wang~\emph{et.~al.} \citep{Wang:2006:RSN:1156422.1156550} propose a NAT traversal algorithm for setting we are using based on
+a changing of a source port, while our algorithms keep source port constant.
 
@@
 
-\subsection{NAT classes}
+\subsection{NAT classes}\label{sec:nat}
 NAT is classified w.r.t. \emph{filtering rule} and \emph{allocation rule} which says what happens after receiving
@@
 
-\paragraph{Full cone NAT.} ~\\
-Allocation table $\mathbb{AT} \subseteq \{ \mathbb{IP}_{in} \times \mathbb{PORT} \} \times \mathbb{PORT}$ holds associations between internal 
-hosts socket and mapped external NAT port. 
-% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
-Filtering rule is:
-\begin{align*}
-& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
-&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}_{in}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
-& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
-\end{align*}
-
-\paragraph{Address restricted cone NAT.} ~\\
-Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex}\} \times \mathbb{PORT}$ 
-holds associations between internal hosts socket, external host IP address and mapped external NAT port. 
-% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
-Filtering rule is:
-\begin{align*}
-& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
-&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
-& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
-\end{align*}
+% MOVED TO APPENDIX
+% \paragraph{Full cone NAT.} ~\\
+% Allocation table $\mathbb{AT} \subseteq \{ \mathbb{IP}_{in} \times \mathbb{PORT} \} \times \mathbb{PORT}$ holds associations between internal 
+% hosts socket and mapped external NAT port. 
+% % Rule is created once the connection from the internal host is initiated to an arbitrary external address.
+% Filtering rule is:
+% \begin{align*}
+% & PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
+% &  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}_{in}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
+% & ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
+% \end{align*}
+%  
+% \paragraph{Address restricted cone NAT.} ~\\
+% Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex}\} \times \mathbb{PORT}$ 
+% holds associations between internal hosts socket, external host IP address and mapped external NAT port. 
+% % Rule is created once the connection from the internal host is initiated to an arbitrary external address.
+% Filtering rule is:
+% \begin{align*}
+% & PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
+% &  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
+% & ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
+% \end{align*}
 
@@
 the new entries in the allocation table are created. For Cone NAT holds that multiple entries in the allocation table 
-can have the same external port assigned, while Symmetric NAT creates new an allocation entries under some circumstances.
+can have the same external port assigned, while Symmetric NAT creates new allocation entries under certain circumstances.
 
@@
 If previously allocated port was $\text{PORT\textsubscript{nat}}$
-the next port will be $\text{PORT\textsubscript{nat}}+\Delta$, where $\Delta=1$ usually. Another type of AF is \emph{random}.
+the next port will be $\text{PORT\textsubscript{nat}}+\Delta$, where $\Delta=1$ usually.
 If the chosen port is already taken, AF iterates until the free one is found.
@@
 \end{align*}
-
 % 
@@
 % \end{align*}
-
-% 
-%We mainly categorize NATs from two perspectives a) 
-% 
-% Hola hop, hola hop, he-ja he-ja hej rup. Hola hop, hola hop, he-ja he-ja hej rup.
-% \begin{compactitem}
-% \item STUN~\citep{rfc5389}
-% \item TURN~\citep{rfc5766}
-% \item Teredo~\citep{rfc4380}
-% \item uPNP~\citep{rfc6970}
-% \item PCP~\citep{rfc6887}
-% \item ICE~\citep{rfc5245}
-% \end{compactitem}
-%
-
-% Properties of NAT assumed in our work:
-% \begin{compactitem}
-% \item symmetric
-% \item port-restricted firewall
-% \item port-sensitive allocation
-% \item incremental port allocation rule
-% \end{compactitem}
 
@@
 \section{UDP hole punching with STUN}
-All algorithms here make use of the same principle, \emph{UDP hole punching}. 
+All algorithms in our paper make use of the same principle, \emph{UDP hole punching}. 
 Internet hosts cannot send packet to the internal host directly. 
@@
 A and B are behind public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$
-respectively. By default an internal host does not known it's external address and NAT type thus technique like 
+respectively. By default an internal host does not known its external address and NAT type thus technique like 
 STUN\footnote{Core idea: publicly available service running on 2 IP addresses and 2 ports 
@@
 it is important to model how NAT's internal state changes over time, i.e., how many external ports 
-were allocated in a given time interval. Thus the workload is important parameter for predicting allocated port.
+were allocated in a given time interval. Thus the workload is important parameter for predicting an allocated port.
 
@@
 Moreover we assume incremental NAT with $\Delta=1$ on both sides. If we sample the NAT 
-state\footnote{One new STUN request to determine current port number, it creates a new NAT allocation record.}
+state\footnote{STUN request to determine current port number, it creates a new NAT allocation record.}
 each $T=10$~milliseconds the NAT state can be modeled as a random process $\{C_i | C_i \in \mathbb{N}, C_i \geq i, i\geq0\}$ where:
@@
 what is common value in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~s on average. 
-If it is higher the NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$.} pretty fast.
+If it is higher the NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$.} quickly.
 
@@
 a connection (i.e., success rate) it proceeds in discrete \emph{steps}. In each step it 
-tries to open a new hole on it's NAT and/or to predict corresponding port allocated on the destination
+tries to open a new hole on the local NAT and/or to predict corresponding port allocated on the destination
 NAT.
@@
 Prior algorithm run it is needed to 
-determine NAT state (UDP hole punching), this is assumed to be done by some technique like STUN. For simplicity we assume the port numbers
+determine NAT state (for UDP hole punching), this is assumed to be done by some technique like STUN. For simplicity we assume the port numbers
 starts from $0$. The length of the silent period is also known to the algorithm prior execution in order to estimate the current state of 
@@
 \paragraph{Time.} 
-The algorithm proceeds in discrete time steps. Algorithm starts in time $t_0 = 0$~ms and each step $i \in \mathcal{S}$
+%The algorithm proceeds in discrete time steps. 
+Algorithm starts in time $t_0 = 0$~ms and each step $i \in \mathcal{S}$
 happens in time $t_i = iT$ for some $T$. If it is not mentioned otherwise, $T = 10$~ms.
@@
 
-\paragraph{Algorithms simulation.}
+\paragraph{Simulation.}
 
@@
 \end{figure}
-
-In figure \ref{fig:alg_eval} is depicted result of a simulation of 
-the algorithms, for different $\lambda$. Figure \ref{fig:alg_steps} depicts number of steps needed 
+Figure \ref{fig:alg_eval} depicts algorithms success rate w.r.t. $\lambda$. 
+Figure \ref{fig:alg_steps} depicts number of steps needed 
 to succeed on average. From figures it is visible Alg. I is more suitable for networks with low 
@@
 
-\begin{itemize}
- \item ToDo: fitting model to the real data from MU network
- \item ToDo: evaluating algorithms on real data from MU network
-\end{itemize}
+% \begin{itemize}
+%  \item ToDo: fitting model to the real data from MU network
+%  \item ToDo: evaluating algorithms on real data from MU network
+% \end{itemize}
 
 \section{Real data modeling}
-Data collected on university network by NetFlow probes were studied to test the correspondence to the Poisson process
+Data collected in our university network by NetFlow probes were studied to test the correspondence to the Poisson process
 w.r.t. newly created connections/allocations on NAT. NAT was simulated in our program using
@@
 If the number of samples is higher, e.g., 500, the time window of studied network traffic is longer and
-in a real world scenario the probability that $\lambda$ is constant in time is low with increasing the time
+in a real world scenario the probability that $\lambda$ is constant in time is low with the increasing length of the time
 window. Due to this time homogeneous model is not strong enough, data is 
@@
 \section{Discussion}
-For simplicity here we assumed the $\lambda$ is the same on both sides, the new challenges 
+For simplicity, we assumed that $\lambda$ is the same on both sides, the new challenges 
 can arise if we allow different parameters on both sides that we haven't studied yet.
@@
 
+To validate our results we implemented Android application using baby-step giant-step algorithm
+to traverse symmetric carrier-grade NAT the mobile phone operator was using. The implementation
+works according to expectations.
+
+\appendix
+\section{NAT classes}\label{sec:classes_appendix}
+Definition of a Full cone NAT and Address restricted cone NAT as given in section \ref{sec:nat}.
+
+\paragraph{Full cone NAT.} ~\\
+Allocation table $\mathbb{AT} \subseteq \{ \mathbb{IP}_{in} \times \mathbb{PORT} \} \times \mathbb{PORT}$ holds associations between internal 
+hosts socket and mapped external NAT port. 
+% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
+Filtering rule is:
+\begin{align*}
+& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
+&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}_{in}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
+& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
+\end{align*}
+
+\paragraph{Address restricted cone NAT.} ~\\
+Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex}\} \times \mathbb{PORT}$ 
+holds associations between internal hosts socket, external host IP address and mapped external NAT port. 
+% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
+Filtering rule is:
+\begin{align*}
+& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
+&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
+& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
+\end{align*}
+
 % -----

### 2013-12-19 paper_lncs.tex
lncs

--- 2013-12-16 paper.tex
@@
-\documentclass[twoside]{article}
+% \documentclass[twoside]{article}
+\documentclass{llncs}
+%\usepackage{llncsdoc}
 
@@
 % Lettrines
-\usepackage{lettrine}
+% \usepackage{lettrine}
 
@@
 % Titling (section/subsection)
-\usepackage{titlesec}
-\renewcommand\thesection{\Roman{section}}
-\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{}
+% \usepackage{titlesec}
+% \renewcommand\thesection{\Roman{section}}
+% \titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{}
 
@@
 % Header/footer
-\usepackage{fancyhdr}
-	\pagestyle{fancy}
-	\fancyhead{}
-	\fancyfoot{}
-	\fancyhead[C]{SPW 2014 $\bullet$ Twenty-second International Workshop on Security Protocols}
-	\fancyfoot[RO,LE]{\thepage}
+% \usepackage{fancyhdr}
+% 	\pagestyle{fancy}
+% 	\fancyhead{}
+% 	\fancyfoot{}
+% 	\fancyhead[C]{SPW 2014 $\bullet$ Twenty-second International Workshop on Security Protocols}
+% 	\fancyfoot[RO,LE]{\thepage}
 % CFP: http://www.wikicfp.com/cfp/servlet/event.showcfp?eventid=33620&copyownerid=44118
@@
 % Math
-\usepackage{amsthm}
+% \usepackage{amsthm}
 \usepackage{amsmath, mathtools}
@@
 \maketitle
-\thispagestyle{fancy}
+% \thispagestyle{fancy}
 
@@
 A presence of such potentially malicious hosts in a communication path is not desired, mainly due to increased cost, poor link quality 
-and some security consequences.
+and security consequences.
 Several solution exist for traversing NAT but symmetric case is still problematic. We propose algorithms using single source port for 
@@
 
-\begin{multicols}{2}
+% \begin{multicols}{2}
 % \lettrine[nindent=0em,lines=3]{L}orem ipsum dolor sit amet.
@@
 Wang~\emph{et.~al.} \citep{Wang:2006:RSN:1156422.1156550} propose a NAT traversal algorithm for setting we are using based on
-a changing of a source port, while our algorithms keep source port constant.
+a changing of a source port, while in our approach we want to explore possibilities with constant source port.
 
@@
 \end{figure}
-Figure \ref{fig:alg_eval} depicts algorithms success rate w.r.t. $\lambda$. 
+Figure \ref{fig:alg_eval} shows algorithms success rate w.r.t. $\lambda$. 
 Figure \ref{fig:alg_steps} depicts number of steps needed 
@@
 and to provide an algorithm with high success rate for networks with low workload that is easy to 
-implement and requires little system resource. 
-Several algorithms solving this problem but there are still some open questions remaining.
+implement and requires little system resources. 
+Several algorithms solving this problem were proposed but there are still some open questions remaining. 
+Our goal was to find the algorithms that maximize success rate using only one source port strategies 
+what can be used in further research, possibly in combination with previous multiple source port strategies.
 
@@
 may be helpful for another researchers interested in this area so we published it in public domain
-hosted on github [URL:WHERE].
+hosted on GitHub [URL:HERE].
 
@@
 
-\end{multicols}
-
+% \end{multicols}
 \end{document}

### 2013-12-31 paper_lncs.tex
finishing paper, introduction

@@
 \usepackage{fixltx2e}
-% \usepackage{wrapfig}
+\usepackage{wrapfig}
 \usepackage{tikz}
@@
 %\usepackage{appendices}
+\usepackage{subcaption}
 
@@
 	}	
-\author{%
-	\large
-	\textsc{Du\v{s}an Klinec} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
-	\normalsize	Faculty of Informatics \\
-	\normalsize	Masaryk University, Brno \\
-	\normalsize	\href{mailto:xklinec@fi.muni.cz}{xklinec@fi.muni.cz}
-	\and
-	\textsc{Vashek Matyáš} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
-	\normalsize	Faculty of Informatics \\
-	\normalsize	Masaryk University, Brno \\
-	\normalsize	\href{mailto:matyas@fi.muni.cz}{matyas@fi.muni.cz}
-	\vspace{-5mm}
-	}
-\date{}
+% \author{%
+% 	\large
+% 	\textsc{Du\v{s}an Klinec} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
+% 	\normalsize	Faculty of Informatics \\
+% 	\normalsize	Masaryk University, Brno \\
+% 	\normalsize	\href{mailto:xklinec@fi.muni.cz}{xklinec@fi.muni.cz}
+% 	\and
+% 	\textsc{Vashek Matyáš} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
+% 	\normalsize	Faculty of Informatics \\
+% 	\normalsize	Masaryk University, Brno \\
+% 	\normalsize	\href{mailto:matyas@fi.muni.cz}{matyas@fi.muni.cz}
+% 	\vspace{-5mm}
+% 	}
+% \date{}
+
+\urldef{\mailsa}\path|xklinec@fi.muni.cz|
+\urldef{\mailsb}\path|matyas@fi.muni.cz|
+\author{Du\v{s}an Klinec \and Vashek Matyáš}
+\institute{Faculty of Informatics, Masaryk University, Brno, Czech Republic,\\
+\email{\{xklinec,matyas\}@fi.muni.cz}}
 
@@
 \section{Introduction}
-Network Address Translators (abbreviated as NAT) has been increasing in popularity in recent years 
-mainly as a mean of slowing down IPv4 address space depletion. On one hand NAT offers
+Network Address Translators (abbreviated as NATs) has been increasing in popularity in recent years, 
+mainly as a mean of slowing down IPv4 address space depletion. On one hand, NAT offers
 the ability to share a public, globally routable IP address among many internal network hosts. On 
-the other hand it breaks basic end-to-end principle of the Internet and thus favors server-client 
+the other hand, it breaks the basic end-to-end principle of the Internet and thus favors server-client 
 communication model for clients connected behind NAT.
@@
 its parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
-between communicating parties is necessary. If P2P connection is not possible to be established, relay 
+between communicating parties is necessary. If P2P connection is not possible, relay 
 servers has to be used (TURN~\citep{rfc5766} protocol is used). The presence 
@@
 (increased latency, jitter, lower bandwidth), relay server is a potential point of failure, 
-has to scale with increasing demands of the network, etc...
+has to scale with increasing demands of the network, etc.
 Thus relay servers require non-negligible resources what makes them expensive for VoIP providers and
@@
 
-From an attacker perspective, relay servers are interesting point where to attack thus from security
-standpoint they could be seen as potential enemies trying to sniff and log all 
+From an attacker perspective, relay servers can be seen as potential enemies trying to sniff and log all 
 communication. NAT (especially symmetric one, used as a carrier-grade border firewall in large networks)
 forces communicating parties to use such central servers
-for communication making monitoring easier for an attacker. Traversal algorithms has to collaborate 
+for communication, making monitoring easier for an attacker. Traversal algorithms has to collaborate 
 with central servers in such a way they create a direct link, not using servers any further. 
@@
 
-Most of the applications use ICE\ignore{~\citep{rfc5245}} nowadays, what is a very powerful toolbox for traversing 
-various types of NAT, however does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for its operation.
+ICE\ignore{~\citep{rfc5245}} is used as a technique to traverse NAT in most applications nowadays, 
+however does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for its operation.
 
 % \section{Related work}
-Some interesting algorithms for traversing symmetric NAT were proposed by Y.~Takeda in \citep{takeda}, but the case both clients are behind symmetric NAT
+Some interesting algorithms for traversing symmetric NAT were proposed by Y.~Takeda in \citep{takeda}, the case where both clients are behind symmetric NAT
 with port sensitive allocation function is not covered. It mentions that sending more packets does not increase a probability of establishing a new connection.
@@
 RFC~5128 \citep{rfc5128} maps current state of the art in NAT traversal, but does not cover our case. 
-The research by Wei~\emph{et.~al.} in \citep{wei}., is devoted to symmetric NAT traversal but their algorithm is using another technique compared
+The research by Wei~\emph{et.~al.} in \citep{wei}., is devoted to symmetric NAT traversal, but their algorithm is using another technique compared
 to our approach. The downside of their algorithm is it requires Time To Live~(TTL) modification of the IP~packet. This is usually considered
@@
 
-Wang~\emph{et.~al.} \citep{Wang:2006:RSN:1156422.1156550} propose a NAT traversal algorithm for setting we are using based on
-a changing of a source port, while in our approach we want to explore possibilities with constant source port.
+Wang~\emph{et.~al.} \citep{Wang:2006:RSN:1156422.1156550} propose a NAT traversal algorithm for scenario where two hosts are
+behind symmetric NATs, based on a changing of a source port, while in our approach we want to explore possibilities with constant source port.
 
-In our work we study the situation where two parties, A,B tries to establish a direct connection, both of them are behind 
-a symmetric NAT (possibly cascade of such NATs) with port sensitive, predictable allocation function. Another scenarios with
+In our work we study the situation where two parties -- A,B -- try to establish a direct connection, both of them are behind 
+a symmetric NAT (possibly cascade of such NATs) with port sensitive, predictable allocation function. Other scenarios with
 predictable allocation function are quite easy to solve and covered in previous research.
@@
 \begin{mydef}
-\begin{align*}
-\mathbb{SC} = \mathbb{IP} \times \mathbb{PORT}
-\end{align*} is a Socket. $\mathbb{IP}$ is set of all possible IP addresses and $\mathbb{PORT}$ 
+% \begin{align*}
+$\mathbb{SC} = \mathbb{IP} \times \mathbb{PORT}$
+% \end{align*}%
+ is a Socket. $\mathbb{IP}$ is set of all possible IP addresses and $\mathbb{PORT}$ 
 is set of all possible ports, basically $\mathbb{PORT} = \{1,2,\dots,65535\}$.
-\begin{align*}
-\mathbb{SP} = \mathbb{IP} \times \mathbb{PORT} \times \mathbb{IP} \times \mathbb{PORT}
-\end{align*} is a Socket Pair.
+% \begin{align*}
+$\mathbb{SP} = \mathbb{IP} \times \mathbb{PORT} \times \mathbb{IP} \times \mathbb{PORT}$
+% \end{align*}%
+ is a Socket Pair.
 \end{mydef}
@@
 Packet received on the external interface has socket pair {(IP\textsubscript{ex},PORT\textsubscript{ex},IP\textsubscript{nat},PORT\textsubscript{nat})}.
-
 Packet received on the internal interface has socket pair {(IP\textsubscript{in},PORT\textsubscript{in},IP\textsubscript{ex},PORT\textsubscript{ex})}.
@@
 
-\paragraph{Port restricted cone NAT.} ~\\
+\paragraph{Port restricted cone NAT} ~\\
 Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex} \times \mathbb{PORT}\} \times \mathbb{PORT}$ 
@@
 
-\paragraph{Symmetric NAT.} ~\\
+\paragraph{Symmetric NAT} ~\\
 Allocation table $\mathbb{AT}$ and filtering rule is the same as in Port restricted Cone NAT. The difference lies in the way 
@@
 the expected value is also most probable value and the algorithm uses maximum likelihood approach.
-
+\newpage
 \subsection{Algorithm IV -- Poisson sampling}
@@
 % 
-\subsection{Silent period in algorithms.} If the length of the silent period is too big
+\subsection{Silent period in algorithms} If the length of the silent period is too big
 the algorithm can have problems with coping with a gap that occurred in NAT ports. Thus 
@@
 
-\begin{figure}[H]
 % \begin{wrapfigure}{r}{0.5\textwidth}
-% \begin{figure*}
-% \begin{center}
-% \leavevmode
-% \centerline
-%{\scalebox{0.40}{\includegraphics[width=\textwidth]{alg_1.pdf}}}
+%  \begin{figure}
+\begin{wrapfigure}{r}{0.5\textwidth}%
+% \wrapfigure{r}{.5\linewidth}{
+     \centering{
+    \begin{subfigure}{0.5\textwidth} 
 	{\scalebox{0.43}{\includegraphics[trim=40 0 0 0]{alg_1.pdf}}}
-% \end{center}
 	\caption{Algorithm evaluation in a simulation, $T=10~ms$. Number of a simulation rounds for one parameter setting is 1000.
@@
 	\label{fig:alg_eval}
-\end{figure}
-% \end{figure*}
+     \end{subfigure}
+%     \end{figure}
 % \end{wrapfigure}
+% \begin{wrapfigure}{r}{0.5\textwidth}    
+     \begin{subfigure}{0.5\textwidth} %\begin{figure}[H]
+	{\scalebox{0.43}{\includegraphics[trim=40 0 0 0]{alg_steps.pdf}}}
+	\caption{Steps of the algorithms needed to success on average w.r.t. $\lambda$. 
+	Algorithms denoted in a same way as in figure \ref{fig:alg_eval}.}
+	\label{fig:alg_steps}
+     \end{subfigure}}%\end{figure}
+    \vspace{-30pt}
+%  \end{figure}
+\end{wrapfigure}
 
-\begin{figure}[H]
-{\scalebox{0.43}{\includegraphics[trim=40 0 0 0]{alg_steps.pdf}}}
-\caption{Steps of the algorithms needed to success on average w.r.t. $\lambda$. Algorithms denoted in a same way as in figure \ref{fig:alg_eval}.}
-\label{fig:alg_steps}
-\end{figure}
 Figure \ref{fig:alg_eval} shows algorithms success rate w.r.t. $\lambda$. 
@@
 
-\section{Real data modeling}
+\subsection{Real data modeling}
 Data collected in our university network by NetFlow probes were studied to test the correspondence to the Poisson process

### 2013-12-31 paper_lncs_vashek.tex
finishing paper, introduction

--- 2013-12-31 paper_lncs.tex
@@
 \author{Du\v{s}an Klinec \and Vashek Matyáš}
-\institute{Faculty of Informatics, Masaryk University, Brno, Czech Republic,\\
+\institute{Faculty of Informatics, Masaryk University, Brno, Czech Republic\\
 \email{\{xklinec,matyas\}@fi.muni.cz}}
@@
 \noindent Network Address Translators cause troubles for VoIP and other P2P services since central servers are needed for communication.
-A presence of such potentially malicious hosts in a communication path is not desired, mainly due to increased cost, poor link quality 
+Presence of such (potentially malicious) hosts in a communication path is not desired, mainly due to increased cost, poor link quality 
 and security consequences.
-Several solution exist for traversing NAT but symmetric case is still problematic. We propose algorithms using single source port for 
+Several solutions exist for traversing NAT, but symmetric case is still problematic. We propose algorithms using single source port for 
 symmetric NAT traversal, each with different properties and applicability. 
-Paper gives a theoretic point of view on the NAT traversal problem by stating problem formally and modeling a network connected to 
-NAT as a random process, in order to evaluate algorithms in a different settings.
-% 
-% ToDo; Will be written as the last part of the paper...
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur magna lorem, tempor sed facilisis vel, porta et turpis.
-Sed et felis a massa dictum posuere.
-Aliquam hendrerit rhoncus ipsum sit amet placerat.
-Duis fringilla est eu arcu mollis faucibus non sit amet eros.
-Vestibulum risus nibh, dapibus vitae laoreet eget, fringilla quis nisl.
-Proin consequat nibh sit amet mauris suscipit tincidunt.
-Sed rutrum, purus nec aliquam faucibus, quam libero venenatis nisi, ut tempor mi sapien vel diam.
-Pellentesque sagittis elit non risus malesuada accumsan.
-Morbi consequat urna.
+Our paper investigates the NAT traversal problem by stating the problem formally and modeling a network connected to 
+NAT as a random process, in order to evaluate algorithms in different settings.
 \end{abstract}
@@
 \section{Introduction}
-Network Address Translators (abbreviated as NATs) has been increasing in popularity in recent years, 
+Network Address Translators (abbreviated as NATs) have been increasing in popularity in recent years, 
 mainly as a mean of slowing down IPv4 address space depletion. On one hand, NAT offers
@@
 
-In many applications is a peer-to-peer (P2P) connection of crucial importance and the quality of a service depends on 
+In many applications, a peer-to-peer (P2P) connection is of crucial importance and the quality of a service depends on 
 its parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
 between communicating parties is necessary. If P2P connection is not possible, relay 
-servers has to be used (TURN~\citep{rfc5766} protocol is used). The presence 
-of another node in the communication path can have negative effects on a communication channel quality 
-(increased latency, jitter, lower bandwidth), relay server is a potential point of failure, 
+servers have to be used (TURN~\citep{rfc5766} protocol is used). The presence 
+of another node in the communication path can have negative effects on the communication channel quality 
+(increased latency, jitter, lower bandwidth), the relay server is a potential point of failure and 
 has to scale with increasing demands of the network, etc.
-Thus relay servers require non-negligible resources what makes them expensive for VoIP providers and
-establishing a direct connection between parties is of high importance.
+Thus relay servers require non-negligible resources and become expensive infrastructure components for VoIP and other service providers where establishing a direct connection between parties is of high importance.
 
-From an attacker perspective, relay servers can be seen as potential enemies trying to sniff and log all 
+From the attacker's perspective, relay servers can be seen as potential allies when trying to sniff and log all 
 communication. NAT (especially symmetric one, used as a carrier-grade border firewall in large networks)
 forces communicating parties to use such central servers
-for communication, making monitoring easier for an attacker. Traversal algorithms has to collaborate 
+for communication, making monitoring easier for an attacker. Traversal algorithms have to collaborate 
 with central servers in such a way they create a direct link, not using servers any further. 
@@
 
-In this paper we are focused on a symmetric NAT with a predictable port allocation function. We propose 
+In this paper, we focused on a symmetric NAT with a predictable port allocation function. We propose 
 new algorithms for establishing direct connection between hosts behind such NATs that are effective and 
-has low system requirements.
+have low system requirements.
 
@@
 
-Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}. But in general 
-it is not widely deployed and it cannot be relied upon in a general case.
+Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}.
+Yet in general, it is not widely deployed and it cannot be relied upon in a general case.
 
 ICE\ignore{~\citep{rfc5245}} is used as a technique to traverse NAT in most applications nowadays, 
-however does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for its operation.
+although it does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for its operation.
 
 % \section{Related work}
-Some interesting algorithms for traversing symmetric NAT were proposed by Y.~Takeda in \citep{takeda}, the case where both clients are behind symmetric NAT
-with port sensitive allocation function is not covered. It mentions that sending more packets does not increase a probability of establishing a new connection.
+Some interesting algorithms for traversing symmetric NAT were proposed by Y.~Takeda in \citep{takeda}, while the case where both clients are behind symmetric NAT
+with a port sensitive allocation function is not covered. Takeda mentions that sending more packets does not increase a probability of establishing a new connection.
 Our motivation is to show that it is actually possible.
@@
 RFC~5128 \citep{rfc5128} maps current state of the art in NAT traversal, but does not cover our case. 
-The research by Wei~\emph{et.~al.} in \citep{wei}., is devoted to symmetric NAT traversal, but their algorithm is using another technique compared
-to our approach. The downside of their algorithm is it requires Time To Live~(TTL) modification of the IP~packet. This is usually considered
+The research by Wei~\emph{et~al.} in \citep{wei}., is devoted to symmetric NAT traversal, but their algorithm is using another technique compared to ours, and the the downside of their algorithm comes with the Time To Live~(TTL) parameter modification of the IP~packet.
+This is usually considered
 as a privileged operation in the operating system and super-user privileges are required for this modification.
 
-Wang~\emph{et.~al.} \citep{Wang:2006:RSN:1156422.1156550} propose a NAT traversal algorithm for scenario where two hosts are
-behind symmetric NATs, based on a changing of a source port, while in our approach we want to explore possibilities with constant source port.
+Wang~\emph{et~al.} \citep{Wang:2006:RSN:1156422.1156550} propose a NAT traversal algorithm for the scenario where two hosts are
+behind symmetric NATs, based on a changing of a source port, while we explore options with a constant source port.
 
-In our work we study the situation where two parties -- A,B -- try to establish a direct connection, both of them are behind 
-a symmetric NAT (possibly cascade of such NATs) with port sensitive, predictable allocation function. Other scenarios with
+In our work we study the situation where two parties -- $A, B$ -- try to establish a direct connection, both of them are behind 
+a symmetric NAT (possibly a cascade of such NATs) with a port sensitive, predictable allocation function. Other scenarios with
 predictable allocation function are quite easy to solve and covered in previous research.
@@
 \section{Network Address Translation}
-The notation used for the rest of the paper follows.
+The notation used for the rest of our paper follows.
 
@@
 % \end{align*}%
- is a Socket. $\mathbb{IP}$ is set of all possible IP addresses and $\mathbb{PORT}$ 
-is set of all possible ports, basically $\mathbb{PORT} = \{1,2,\dots,65535\}$.
+ is a Socket. $\mathbb{IP}$ is the set of all possible IP addresses and $\mathbb{PORT}$ 
+is the set of all possible ports, basically $\mathbb{PORT} = \{1,2,\dots,65535\}$.
 % \begin{align*}
@@
 \paragraph{NAT and communication.}
-NAT connects two networks, denoted as an internal and an external, connected to an internal and an external NAT interface respectively.
-Hosts has addresses $\text{IP\textsubscript{in}} \in \mathbb{IP}_{in} \subseteq \mathbb{IP}$ 
-and $\text{IP\textsubscript{ex}} \in \mathbb{IP}_{ex} \subseteq \mathbb{IP}$ in the internal and the external network
-respectively, $\mathbb{IP}_{in} \cap \mathbb{IP}_{ex} = \emptyset$.
-If it is not said otherwise the external NAT address\footnote{We assume NAT has only one external address for simplicity.} is denoted as 
+NAT connects two networks, denoted as {\em internal} and {\em external}, connected to an internal and an external NAT interface, respectively.
+Hosts have addresses $\text{IP\textsubscript{in}} \in \mathbb{IP}_{in} \subseteq \mathbb{IP}$ 
+and $\text{IP\textsubscript{ex}} \in \mathbb{IP}_{ex} \subseteq \mathbb{IP}$ in the internal and the external network,
+respectively, where $\mathbb{IP}_{in} \cap \mathbb{IP}_{ex} = \emptyset$.
+If it is not stated otherwise the external NAT address\footnote{We assume NAT has only one external address for simplicity.} is denoted as 
 $\text{IP\textsubscript{nat}} \in \mathbb{IP}_{ex}$.
 
-Packet received on the external interface has socket pair {(IP\textsubscript{ex},PORT\textsubscript{ex},IP\textsubscript{nat},PORT\textsubscript{nat})}.
-Packet received on the internal interface has socket pair {(IP\textsubscript{in},PORT\textsubscript{in},IP\textsubscript{ex},PORT\textsubscript{ex})}.
+Packet received on the external interface has the socket pair {(IP\textsubscript{ex},PORT\textsubscript{ex},IP\textsubscript{nat},PORT\textsubscript{nat})}.
+Packet received on the internal interface has the socket pair {(IP\textsubscript{in},PORT\textsubscript{in},IP\textsubscript{ex},PORT\textsubscript{ex})}.
 
@@
 \subsection{NAT classes}\label{sec:nat}
-NAT is classified w.r.t. \emph{filtering rule} and \emph{allocation rule} which says what happens after receiving
-packet from an external and an internal network respectively.
+NAT is classified w.r.t. the \emph{filtering rule} and the \emph{allocation rule}, which state what happens after receiving
+packet from an external and an internal network, respectively.
 
-Common NAT categorization is according to~\citep{rfc3489}: a) Full Cone
-b) IP Restricted Cone, c) Port Restricted Cone, d) Symmetric.
+Common NAT categorization is according to~\citep{rfc3489}: a) Full Cone;
+b) IP Restricted Cone; c) Port Restricted Cone; d) Symmetric.
 
@@
 
-\paragraph{Port restricted cone NAT} ~\\
+\paragraph{Port Restricted Cone NAT} ~\\
 Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex} \times \mathbb{PORT}\} \times \mathbb{PORT}$ 
-holds associations between internal hosts socket, external host socket and mapped external NAT port. 
+holds associations between the internal host socket, external host socket and mapped external NAT port. 
 % Rule is created once the connection from the internal host is initiated to an arbitrary external address.
-Filtering rule is:
+The filtering rule is:
 \begin{align*}
@@
 \paragraph{Symmetric NAT} ~\\
-Allocation table $\mathbb{AT}$ and filtering rule is the same as in Port restricted Cone NAT. The difference lies in the way 
-the new entries in the allocation table are created. For Cone NAT holds that multiple entries in the allocation table 
-can have the same external port assigned, while Symmetric NAT creates new allocation entries under certain circumstances.
+Allocation table $\mathbb{AT}$ and filtering rule are the same as in the Port Restricted Cone NAT. The difference lies in the way 
+the new entries in the allocation table are created. For the Cone NAT holds that multiple entries in the allocation table 
+can have the same external port assigned, while the Symmetric NAT creates new allocation entries under certain circumstances.
 
 \emph{Allocation function (AF)} defines port allocation rules for new connections (not in $\mathbb{AT}$). Assume $\mathbb{AT}=\emptyset$, for simplicity. 
-AF can be either \emph{predictable} or \emph{random}. In our setting predictable AF is assumed, e.g., \emph{incremental}. 
-If previously allocated port was $\text{PORT\textsubscript{nat}}$
-the next port will be $\text{PORT\textsubscript{nat}}+\Delta$, where $\Delta=1$ usually.
-If the chosen port is already taken, AF iterates until the free one is found.
+AF can be either \emph{predictable} or \emph{random}. In our setting, a predictable AF is assumed, e.g., \emph{incremental}. 
+If the previously allocated port was $\text{PORT\textsubscript{nat}}$ then
+the next port will be $\text{PORT\textsubscript{nat}}+\Delta$, where typically $\Delta=1$.
+If the chosen port is already taken, AF iterates until a free one is found.
 
@@
 All algorithms in our paper make use of the same principle, \emph{UDP hole punching}. 
-Internet hosts cannot send packet to the internal host directly. 
+Internet hosts cannot send packets to the internal host directly. 
 Communication has to be initiated from the inside\footnote{Not taking
-pre-set port forwarding, DMZ, uPnP, etc... into consideration. }, what is problematic if two nodes behind 
-NAT wants to communicate.
+pre-set port forwarding, DMZ, uPnP, etc. into consideration. }, which is problematic if two nodes behind 
+NAT want to communicate.
 
-Assume two parties, A and B with addresses IP\textsuperscript{A} and IP\textsuperscript{B} respectively,
-A and B are behind public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$
-respectively. By default an internal host does not known its external address and NAT type thus technique like 
+Assume two parties, A and B with addresses IP\textsuperscript{A} and IP\textsuperscript{B}, respectively.
+A and B are behind a public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$,
+respectively. By default an internal host does not know its external address and NAT type, thus a technique like 
 STUN\footnote{Core idea: publicly available service running on 2 IP addresses and 2 ports 
-answering questions like: what is my IP address and port? Change IP address of response if I'll get it, etc...} 
-is used to learn it. In case of incremental NAT it is also needed to determine $\Delta$ and the next port
+answering questions like: what is my IP address and port? Change IP address of response if Il get it, etc.} 
+is used to learn it. In case of incremental NAT it is also necessary to determine $\Delta$ and the next port
 that is likely to be allocated for a new connection $\text{PORT}^{\text{A}}_{\text{nat}}$ for A and $\text{PORT}^{\text{B}}_{\text{nat}}$
@@
 In order to evaluate algorithms in multiple environments with different parameters (workload, number of clients, etc.)
-we decided to model network behavior with stochastic process. In particular we choose a Poisson process, $\{N(t), t\geq0\}$ 
+we decided to model network behavior with a stochastic process. In particular, we chose a Poisson process, $\{N(t), t\geq0\}$ 
 that is used in queueing theory to model arrivals of requests to the server queue~\citep{Nelson:1995:PSP:207382}.
@@
 The core idea of NAT traversal is to predict the next allocated external port by NAT, thus from this perspective 
-it is important to model how NAT's internal state changes over time, i.e., how many external ports 
-were allocated in a given time interval. Thus the workload is important parameter for predicting an allocated port.
+it is important to model how the NAT internal state changes over time, i.e., how many external ports 
+have been allocated in a given time interval. Thus the workload is an important parameter for predicting the allocated port.
 
-We model internal network connected to NAT w.r.t. newly created connections (new port allocation) as a 
+We model the internal network connected to NAT w.r.t. newly created connections (new port allocation) as a 
 time homogeneous Poisson process $Po(\lambda)$. By setting $\lambda$ we can model 
-different workload of the network and evaluate algorithms with respect to the various parameters e.g., 
-probability of an establishing connection.
+different workloads of the network and evaluate algorithms w.r.t. various parameters, e.g., 
+probability of establishing a connection.
 
@@
 
-Moreover we assume incremental NAT with $\Delta=1$ on both sides. If we sample the NAT 
-state\footnote{STUN request to determine current port number, it creates a new NAT allocation record.}
+Moreover, we assume an incremental NAT with $\Delta=1$ on both sides. If we sample the NAT 
+state\footnote{STUN requests to determine current port number, it creates a new NAT allocation record.}
 each $T=10$~milliseconds the NAT state can be modeled as a random process $\{C_i | C_i \in \mathbb{N}, C_i \geq i, i\geq0\}$ where:
@@
 
-\paragraph{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected in this time interval.} is assumed to be 3~minutes, 
-what is common value in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~s on average. 
-If it is higher the NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$.} quickly.
+\paragraph{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected within this time interval.} is assumed to be 3~minutes, 
+which is a typical setting in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~s on average. 
+If it is higher then NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$.} quickly.
 
 \section{Generic algorithm structure}
-Algorithms we propose have a common structure. In order to maximize probability of an establishing
+Algorithms that we propose have a common structure. In order to maximize the probability of establishing
 a connection (i.e., success rate) it proceeds in discrete \emph{steps}. In each step it 
-tries to open a new hole on the local NAT and/or to predict corresponding port allocated on the destination
+tries to open a new hole in the local NAT and/or to predict the corresponding port allocated in the destination
 NAT.
@@
 \begin{compactitem}
-\item $\mathcal{U}=\{0,1\}$ is set of communicating parties.
+\item $\mathcal{U}=\{0,1\}$ is the set of communicating parties.
 \item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$ is a step of the algorithm.
-\item $\mathcal{M}$ is set of vectors of parameters of the environment (i.e., network model 
+\item $\mathcal{M}$ is the set of vectors of parameters of the environment (i.e., network model 
 parameters and additional information obtained in initial phase). 
-In our setting $\mathcal{M} = \{ (\lambda, t_s) | \lambda~>~0, t\geq~0 \}$. Value $t_s$ is length of the silent period. %For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
+In our setting $\mathcal{M} = \{ (\lambda, t_s) | \lambda~>~0, t\geq~0 \}$. Value $t_s$ denotes the length of the silent period. %For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
 \item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
@@
 \end{mydef}
-If the third parameter is not important for the algorithm it can be ignored in the notation.
+In case $\mathcal{M}$ is not not important for the algorithm it can be ignored in the notation.
+%If the third parameter is not important for the algorithm it can be ignored in the notation.
 
 Prior algorithm run it is needed to 
-determine NAT state (for UDP hole punching), this is assumed to be done by some technique like STUN. For simplicity we assume the port numbers
-starts from $0$. The length of the silent period is also known to the algorithm prior execution in order to estimate the current state of 
-the NAT, e.g., difference between the last measurement and just before algorithm execution.
+determine the NAT state (for UDP hole punching), this is assumed to be done by some technique like STUN. For simplicity, we assume that port numbers
+start from $0$. The length of the silent period is also known to the algorithm prior execution in order to estimate the current state of 
+the NAT, e.g., difference between the last measurement and just before actual algorithm execution.
 
-\paragraph{Intuition.} 
+\paragraph{Notation.}
 Line $A(0,1) = (2,3)$ means party $0$ in step $1$ sent packet from source port $2$ to destination port $3$
@@
 Algorithm starts in time $t_0 = 0$~ms and each step $i \in \mathcal{S}$
-happens in time $t_i = iT$ for some $T$. If it is not mentioned otherwise, $T = 10$~ms.
+happens in time $t_i = iT$ for some $T$. If not mentioned otherwise, $T = 10$~ms.
 
@@
 
-\paragraph{Algorithm log \& successful run.} IP addresses of the hosts and NAT are known and fixed prior algorithm run thus omitted
-from the notation for simplicity. For the following text is assumed maximum number of steps of the algorithm $s_{max}=1000$.
+\paragraph{Algorithm log \& successful run.} IP addresses of the hosts and NAT are known and fixed prior to the algorithm run, and are thus omitted from the notation for simplicity.
+For the following text we assume the maximum number of steps of the algorithm $s_{max}=1000$.
 
@@
  \item $(p_1, p_2) = \mathcal{A}(0, s, (\lambda, t_s)) $ are generated by the algorithm $\mathcal{A}$ for some 
-       parameters $(\lambda, t_s)$ in some step $s$ of the algorithm $\mathcal{A}$.
+       parameters $(\lambda, t_s)$ in some step $s$ of the algorithm $\mathcal{A}$;
  \item $p_3$ is an allocated port by $NAT^A$ for request $(p_1, p_2)$ in a given step $s$. In our network model
-       $p_3=C_s=C_{s-1} + 1 + X, X~\sim~Po(\lambda T)$.
- \item Algorithm log for host $B$ is defined analogically.
- \item Furthermore from our setting the following formula holds:\footnote{From definition of a port sensitive symmetric NAT.}
+       $p_3=C_s=C_{s-1} + 1 + X, X~\sim~Po(\lambda T)$;
+ \item algorithm log for host $B$ is defined analogically;
+ \item furthermore from our setting the following formula holds:\footnote{From definition of a port sensitive symmetric NAT.}
 \end{compactitem} %
@@
 \label{def:match}
-Algorithm $\mathcal{A}$ run is successful, i.e., established a direct connection between $A,B$ if\footnote{Derived from UDP hole punching.}:
+Algorithm $\mathcal{A}$ run is successful, i.e., established a direct connection between $A,B$ if~\footnote{Derived from UDP hole punching.}:
 \begin{align*}
 & MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}}) \Leftrightarrow \exists(p_1, p_2, p_3) \in \mathcal{L}^{A}_{\mathcal{A}}: \\
-& \exists(p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{B}_{\mathcal{A}}: p_2 = p_3^{\prime} \wedge p_2^{\prime} = p_3
+& \exists(p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{B}_{\mathcal{A}}: p_2 = p_3^{\prime} \wedge p_2^{\prime} = p_3.
 %
@@
 \end{mydef}
-The task is to find an algorithm $\mathcal{A}$ such that the probability the $MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}})$
+The task is to find an algorithm $\mathcal{A}$ such that the probability $MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}})$
 holds is maximized for minimal number of steps needed.
@@
 We propose the algorithm with high success rate for small network loads, i.e.,
-$\lambda \leq 0.03, T=10$~ms. When error becomes larger,
-algorithm step is not fast enough to compensate errors and thus fails for larger $\lambda$.
+$\lambda \leq 0.03, T=10$~ms. When the error becomes larger, the
+algorithm step is not fast enough to compensate errors and thus fails for a larger $\lambda$.
 
@@
 
-Intuitively the source port is kept constant, one side of the protocol scans ports sequentially, another
+Intuitively, the source port is kept constant, one side of the protocol scans ports sequentially, another
 side of the protocol scans ports with step 2.
@@
 \mathcal{A}(u, s) = (s, \Delta_u)
-\] for some constants $\Delta_0, \Delta_1$ based on the current network workload determined in initial phase
-of the protocol.
+\] for some constants $\Delta_0, \Delta_1$ based on the current network workload determined in initial phase of the protocol.
 \end{mydef}
@@
 
-In step $i$ algorithm tries the expected value for $C_i$. Since the $C_i$ has unimodal distribution
-the expected value is also most probable value and the algorithm uses maximum likelihood approach.
+In step $i$ the algorithm tries the expected value for $C_i$. Since the $C_i$ has unimodal distribution,
+the expected value is also the most probable value and the algorithm uses maximum likelihood approach.
 \newpage
@@
 % 
-\subsection{Silent period in algorithms} If the length of the silent period is too big
-the algorithm can have problems with coping with a gap that occurred in NAT ports. Thus 
-Algorithms III and IV estimates the starting offset caused by the silent period as an
+\subsection{Silent period in algorithms} If the length of the silent period is too big then
+the algorithm can face problems with a gap that occurred in NAT ports. Thus 
+Algorithms III and IV estimate the starting offset caused by the silent period as an
 expected value $E[Po(\lambda \cdot t_1)] = \lambda\cdot t_1$.
@@
 
-\paragraph{Algorithm I and II comparison.} 
-Algorithm I has following benefits over II:
+\paragraph{Comparison of Algorithms I and II.} 
+Algorithm I has the following benefits over II:
 \begin{compactitem}
  \item Works with high success rate for $\lambda \in [0, 0.035] \sim 35$ new connections in 1~sec on average.
- \item Can be stopped as soon as connection is established, overhead is proportional to~$\lambda$.
+ \item Can be stopped as soon as the connection is established, overhead is proportional to~$\lambda$.
  \item Single source port is used, only one listening thread is needed to implement
-it\footnote{NAT is abused to multiplex multiple connections to one source port.} compared to the Alg. II.
-Due to this fact a practical implementation is simple and lightweight w.r.t. needed system resources.
- \item Is $\lambda$-invariant. Provided network load is low no additional measurements are needed.
+it\footnote{NAT is abused to multiplex multiple connections to one source port.} compared to Alg. II.
+Due to this fact a practical implementation is simple and lightweight w.r.t. required system resources.
+ \item Is $\lambda$-invariant. Provided network load is low and no additional measurements are needed.
 \end{compactitem}
@@
 	A, B, C correspond to the Algorithm I, II and III respectively.
-	D, E correspond to the Algorithm IV with optimized $C(\lambda T)$ by an exhaustive search in D case, 
-	in E case $C(\lambda T)$ is approximated by a polynomial.}
+	D, E correspond to the Algorithm IV with optimized $C(\lambda T)$ by an exhaustive search in case D, 
+	in case E is then$C(\lambda T)$ is approximated by a polynomial.}
 	\label{fig:alg_eval}
@@
 workload since has $100\%$ success rate for $\lambda \in [0, 0.035]$. Outside this interval 
-Alg. I is not working and another solution has to be applied. The problem is new connections are
+Alg. I is not working and another solution has to be applied. The problem comes as new connections are
 created faster than algorithm progresses.
@@
 
-Alg. IV has high success rate for 
-noisy networks but is very sensitive on a $\lambda$ estimation what decreases practical usability 
+Alg. IV has a high success rate for 
+noisy networks, but is very sensitive for the $\lambda$ estimation -- this decreases practical usability 
 if $\lambda$ changes quickly over time (non-homogeneous Poisson process). Closed form formula for $C(\lambda T)$ 
@@
 
-Regarding data from set 13, during activity peak during the day, we conclude that if a number of samples 
-is lower, e.g., 100, the process of a creating a new connection on NAT can 
+Regarding data from set 13, during the activity peak during the day, we conclude that if a number of samples 
+is lower, e.g., 100, the process of creating a new connection on NAT can 
 be modeled as a time homogeneous Poisson process successfully (hypothesis rejected in $5.191\%$ of cases).
-The time window the network was monitored in this case is $T \cdot 100 = 1000~ms$.
+The time window in which the network was monitored in this case is $T \cdot 100 = 1000~ms$.
 
 If the number of samples is higher, e.g., 500, the time window of studied network traffic is longer and
-in a real world scenario the probability that $\lambda$ is constant in time is low with the increasing length of the time
-window. Due to this time homogeneous model is not strong enough, data is 
-overdispersed what causes failure to fit Poisson distribution (hypothesis rejected in $20.368\%$ of cases). As suggested
-in literature, Negative Binomial model fits better for overdispersed\footnote{The data overdispersion may be caused by dependent events on the network e.g., time-triggered updates downloading, etc...}
+in a real world scenario the probability that $\lambda$ is constant in time is lower with the increasing length of the time
+window. Due to this the homogeneous model is not strong enough, data is 
+overdispersed and this causes failure to fit Poisson distribution (hypothesis rejected in $20.368\%$ of cases). As suggested
+in literature, Negative Binomial model fits better for overdispersed\footnote{The data overdispersion may be caused by dependent events on the network, e.g., by time-triggered updates downloading.}
 data (hypothesis rejected in $2.829\%$ of cases). 
-The open question is how to make use of this model in NAT traversal.
+The open question is how to make use of this model in the NAT traversal.
 
-For 100 samples the match with the model is satisfactory. Provided algorithm is able to run completely (including learning $\lambda$),
-the proposed algorithms should work like in a simulation.
+For 100 samples the match with the model is satisfactory. Provided the algorithm is able to run completely (including learning $\lambda$), the proposed algorithms should work like in a simulation.
 
-Regarding data from set 03, we can conclude that lower network workload causes more stable mean of sampled ports (MLE for Poisson distribution),
-thus the Poisson distribution hypothesis is not rejected even in longer time intervals for an estimated $\lambda$.
+Regarding data from set 03, we can conclude that a lower network workload causes more stable mean of sampled ports (MLE for Poisson distribution), thus the Poisson distribution hypothesis is not rejected even in longer time intervals for an estimated $\lambda$.
 
-\section{Discussion}
+\section{Discussion and implementation}
 For simplicity, we assumed that $\lambda$ is the same on both sides, the new challenges 
-can arise if we allow different parameters on both sides that we haven't studied yet.
+can arise if we allow different parameters on both sides that we have not studied yet.
 
@@
 Poisson process, but this method requires setting a coefficient function $C(\lambda T)$ for which 
-we don't have exact analytic explanation. This method is also very sensitive to parameter fluctuations
+we do not have exact analytic explanation. This method is also very sensitive to parameter fluctuations
 and thus not very practical for a real world networks.
@@
 In our setting we assume $T=10~ms$ in order to minimize the speed of the evolution of a Poisson process.
-Such short sending interval can cause some packets may get dropped on the path to the destination. But the 
-important is whether they reach the local NAT in order to create a mapping in allocation table. As proposed 
-in \citep{Wang:2006:RSN:1156422.1156550} it is possible to run the algorithm with $T=10~ms$ to create a mapping
+Such a short sending interval may cause some packets to get dropped on the path to the destination. Yet the 
+important issue for us is whether they reach the local NAT in order to create a mapping in the allocation table. As proposed 
+in \citep{Wang:2006:RSN:1156422.1156550}, it is possible to run the algorithm with $T=10~ms$ to create a mapping
 and then re-run the algorithm with higher $T$, e.g., $T=100~ms$ in order to deliver packets to the destination
@@
 Success rate of the algorithm strongly relies on accurate measurements of the network 
-properties the client resides in. They can be measured prior connection establishment 
-from the point the client connects to the network, as is usually done in ICE and gather 
-more information, build better network model and so on in this way.
+properties the client resides in. They can be measured prior to the connection establishment 
+from the point the client connects to the network, as is usually done in ICE, and to gather 
+more information, then to build better network model and to carry on in with a better performance afterwards.
 
@@
 
-Another problematic thing is if this algorithm is followed by a multiple host pairs on
+Another problematic issue comes when this algorithm is followed by multiple host pairs on
 the same network in the same time. They would artificially increase the network workload
-causing it wouldn't follow Poisson distribution rendering our algorithms as not usable in
+causing it would not follow Poisson distribution rendering our algorithms not usable in
 this case.
-
-\section{Conclusion}
-The aim of this work was to study problem of the NAT traversal from statistical point of view,
-and to provide an algorithm with high success rate for networks with low workload that is easy to 
-implement and requires little system resources. 
-Several algorithms solving this problem were proposed but there are still some open questions remaining. 
-Our goal was to find the algorithms that maximize success rate using only one source port strategies 
-what can be used in further research, possibly in combination with previous multiple source port strategies.
-
-We modeled the network workload as a Poisson process what we verified on real data by hypothesis testing.
 
@@
 
-To validate our results we implemented Android application using baby-step giant-step algorithm
-to traverse symmetric carrier-grade NAT the mobile phone operator was using. The implementation
+To validate our results, we implemented Android application using our baby-step giant-step algorithm
+to traverse symmetric carrier-grade NAT that our mobile phone operator was using. The implementation
 works according to expectations.
+
+\section{Conclusion}
+The aim of our work was to study the problem of NAT traversal from a statistical point of view,
+and to provide an algorithm (or algorithms) with a high success rate for networks with a low workload that would be easy to 
+implement and would require little system resources.
+
+We propose several algorithms solving this problem, yet with some open questions remaining. 
+Our goal was to find the algorithms that maximize success rate using only one source port strategy, possibly using this
+in further research, perhaps in combination with previous multiple source port strategies.
+
+We modeled the network workload as a Poisson process and we verified our approach on real data by hypothesis testing.
+Our testing application was published in public domain at GitHub, and we also validated our results by traversing symmetric 
+carrier-grade NAT of our mobile phone operator.
 

### 2014-07-15 NAT_paper_ACM_short.tex
ACM version for SIN2014

--- 2013-12-31 paper_lncs_vashek.tex
@@
-% \documentclass[twoside]{article}
-\documentclass{llncs}
-%\usepackage{llncsdoc}
+\documentclass{acm_proc_article-sp}
+% http://www.acm.org/sigs/publications/proceedings-templates
 
@@
 \usepackage[utf8]{inputenc}
-\linespread{1.05} % Palatino needs more space between lines
+% \linespread{1.05} % Palatino needs more space between lines
 \usepackage{microtype}
@@
 % Page layout
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
+% \usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
 \usepackage[font=it]{caption}
 \usepackage{paralist}
-\usepackage{multicol}
+% \usepackage{multicol}
 
@@
 % Abstract
-\usepackage{abstract}
-	\renewcommand{\abstractnamefont}{\normalfont\bfseries}
-	\renewcommand{\abstracttextfont}{\normalfont\small\itshape}
-
-% ------
-% Titling (section/subsection)
-% \usepackage{titlesec}
-% \renewcommand\thesection{\Roman{section}}
-% \titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{}
-
-% ------
-% Header/footer
-% \usepackage{fancyhdr}
-% 	\pagestyle{fancy}
-% 	\fancyhead{}
-% 	\fancyfoot{}
-% 	\fancyhead[C]{SPW 2014 $\bullet$ Twenty-second International Workshop on Security Protocols}
-% 	\fancyfoot[RO,LE]{\thepage}
-% CFP: http://www.wikicfp.com/cfp/servlet/event.showcfp?eventid=33620&copyownerid=44118
+% \usepackage{abstract}
+% 	\renewcommand{\abstractnamefont}{\normalfont\bfseries}
+% 	\renewcommand{\abstracttextfont}{\normalfont\small\itshape}
 
@@
 \usepackage{wrapfig}
-\usepackage{tikz}
+% \usepackage{tikz}
 \usepackage{float}
@@
 
-% ------
-% Maketitle metadata
-\title{\vspace{-15mm}%
-	\fontsize{24pt}{10pt}\selectfont
-	\textbf{Traversing symmetric NAT with predictable port allocation function}
-	}	
-% \author{%
-% 	\large
-% 	\textsc{Du\v{s}an Klinec} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
-% 	\normalsize	Faculty of Informatics \\
-% 	\normalsize	Masaryk University, Brno \\
-% 	\normalsize	\href{mailto:xklinec@fi.muni.cz}{xklinec@fi.muni.cz}
-% 	\and
-% 	\textsc{Vashek Matyáš} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
-% 	\normalsize	Faculty of Informatics \\
-% 	\normalsize	Masaryk University, Brno \\
-% 	\normalsize	\href{mailto:matyas@fi.muni.cz}{matyas@fi.muni.cz}
-% 	\vspace{-5mm}
-% 	}
-% \date{}
-
-\urldef{\mailsa}\path|xklinec@fi.muni.cz|
-\urldef{\mailsb}\path|matyas@fi.muni.cz|
-\author{Du\v{s}an Klinec \and Vashek Matyáš}
-\institute{Faculty of Informatics, Masaryk University, Brno, Czech Republic\\
-\email{\{xklinec,matyas\}@fi.muni.cz}}
-
 %%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
+\title{Traversing symmetric NAT with predictable port allocation}
+
+\numberofauthors{1}
+\author{
+% 1st. author
+\alignauthor
+Du\v{s}an Klinec, Vashek Maty\'{a}\v{s}\\
+       \affaddr{Faculty of Informatics, Masaryk University}\\
+       \affaddr{Brno, Czech Republic}\\
+       \email{\{xklinec,matyas\}@fi.muni.cz}
+% % 2nd. author
+% \alignauthor
+% G.K.M. Tobin\titlenote{The secretary disavows
+% any knowledge of this author's actions.}\\
+%        \affaddr{Institute for Clarity in Documentation}\\
+%        \affaddr{P.O. Box 1212}\\
+%        \affaddr{Dublin, Ohio 43017-6221}\\
+%        \email{webmaster@marysville-ohio.com}
+% % 3rd. author
+}
 
 \maketitle
-% \thispagestyle{fancy}
-
 \begin{abstract}
-\noindent Network Address Translators cause troubles for VoIP and other P2P services since central servers are needed for communication.
-Presence of such (potentially malicious) hosts in a communication path is not desired, mainly due to increased cost, poor link quality 
-and security consequences.
-Several solutions exist for traversing NAT, but symmetric case is still problematic. We propose algorithms using single source port for 
-symmetric NAT traversal, each with different properties and applicability. 
-Our paper investigates the NAT traversal problem by stating the problem formally and modeling a network connected to 
-NAT as a random process, in order to evaluate algorithms in different settings.
+\noindent Network Address Translators often cause trouble for VoIP and other P2P services since central servers are needed for communication.
+The presence of such potentially malicious hosts in a communication path is not desired, mainly due to security consequences, poor link quality 
+and increased cost.
+Several solutions exist for traversing NAT, but a symmetric one is still problematic. We propose algorithms using a single source port for 
+symmetric NAT traversal. Each with different properties and applicability. 
+% Our paper investigates the NAT traversal problem by stating the problem formally and modeling a network connected to 
+% NAT as a random process, in order to evaluate algorithms in different settings.
 \end{abstract}
 
-% \begin{multicols}{2}
-% \lettrine[nindent=0em,lines=3]{L}orem ipsum dolor sit amet.
-
 \section{Introduction}
-Network Address Translators (abbreviated as NATs) have been increasing in popularity in recent years, 
-mainly as a mean of slowing down IPv4 address space depletion. On one hand, NAT offers
+Network Address Translators (NATs) have been increasing in popularity in recent years, 
+mainly as a means of slowing down IPv4 address space depletion. On one hand, NAT offers
 the ability to share a public, globally routable IP address among many internal network hosts. On 
@@
 its parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
-between communicating parties is necessary. If P2P connection is not possible, relay 
+between communicating parties is necessary. If P2P connection is not possible to establish, relay 
 servers have to be used (TURN~\citep{rfc5766} protocol is used). The presence 
 of another node in the communication path can have negative effects on the communication channel quality 
-(increased latency, jitter, lower bandwidth), the relay server is a potential point of failure and 
+such as increased latency, jitter, lower bandwidth. The relay server is a potential single point of failure and 
 has to scale with increasing demands of the network, etc.
-Thus relay servers require non-negligible resources and become expensive infrastructure components for VoIP and other service providers where establishing a direct connection between parties is of high importance.
+% Thus relay servers require non-negligible resources and become expensive infrastructure components for VoIP and other service 
+% providers where establishing a direct connection between parties is of high importance.
 
-From the attacker's perspective, relay servers can be seen as potential allies when trying to sniff and log all 
-communication. NAT (especially symmetric one, used as a carrier-grade border firewall in large networks)
-forces communicating parties to use such central servers
-for communication, making monitoring easier for an attacker. Traversal algorithms have to collaborate 
-with central servers in such a way they create a direct link, not using servers any further. 
-% 
-% some way. They use it to create a direct link, not using servers any further.
-
-In this paper, we focused on a symmetric NAT with a predictable port allocation function. We propose 
-new algorithms for establishing direct connection between hosts behind such NATs that are effective and 
-have low system requirements.
+In this paper, we focused on a symmetric NAT with a predictable port allocation function. 
+The contributions of this paper are as follows:%
+% \begin{compactitem}
+% \item 
+a) we propose new effective algorithms for establishing a direct connection between hosts behind such NATs; %
+% \item 
+b) we demonstrate the implementations of these algorithms; %, together with a simulation based on a theoretical approach from a queueing theory. These implementations we provide in public domain.%
+% \item 
+c) we tested our algorithms using data from a real network; % (of our mobile phone operator).%
+% \item 
+d) last but not least, we provide a formalization of the NAT traversal problem.
+% \end{compactitem}
+% We propose 
+% new algorithms for establishing direct connection between hosts behind such NATs that are effective and 
+% have low system requirements.
 
 \section{Existing techniques}
-Many different solutions for traversing non-symmetric NATs exist e.g., NAT traversal in Internet Key Exchange~\citep{rfc3947} or 
+Many different solutions for traversing non-symmetric NATs exist, e.g., NAT traversal in Internet Key Exchange~\citep{rfc3947} or 
 Interactive Connectivity Establishment~(ICE)~\citep{rfc5245}. %Terredo\ignore{tunneling}~\citep{rfc4380}.
 
-Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}.
-Yet in general, it is not widely deployed and it cannot be relied upon in a general case.
-
-ICE\ignore{~\citep{rfc5245}} is used as a technique to traverse NAT in most applications nowadays, 
+Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}. However, in general, 
+it is not widely deployed and it cannot be relied upon in a general case. 
+Another solution is ICE\ignore{~\citep{rfc5245}}. It is used as a technique to traverse NAT in most applications nowadays 
 although it does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for its operation.
@@
 % \section{Related work}
-Some interesting algorithms for traversing symmetric NAT were proposed by Y.~Takeda in \citep{takeda}, while the case where both clients are behind symmetric NAT
-with a port sensitive allocation function is not covered. Takeda mentions that sending more packets does not increase a probability of establishing a new connection.
-Our motivation is to show that it is actually possible.
+Interesting algorithms for traversing symmetric NAT were proposed by Y.~Takeda in \citep{takeda}, although does not cover the case with both clients using symmetric NAT
+with a port sensitive allocation function. Takeda mentions that sending more packets does not increase a probability of establishing a new connection.
+We show that it is actually possible. Wang~\emph{et~al.} \citep{Wang:2006:RSN:1156422.1156550} propose a NAT traversal algorithm for the scenario where two hosts are
+behind symmetric NATs, based on changing of a source port, while we explore options with a constant source port.
 
 RFC~5128 \citep{rfc5128} maps current state of the art in NAT traversal, but does not cover our case. 
-The research by Wei~\emph{et~al.} in \citep{wei}., is devoted to symmetric NAT traversal, but their algorithm is using another technique compared to ours, and the the downside of their algorithm comes with the Time To Live~(TTL) parameter modification of the IP~packet.
-This is usually considered
+The research by Wei~\emph{et~al.} in \citep{wei} is devoted to symmetric NAT traversal, but their algorithm is using a different from ours.
+Disadvantage of their approach comes with the Time To Live~(TTL) parameter modification of the IP~packet. This is usually considered
 as a privileged operation in the operating system and super-user privileges are required for this modification.
 
-Wang~\emph{et~al.} \citep{Wang:2006:RSN:1156422.1156550} propose a NAT traversal algorithm for the scenario where two hosts are
-behind symmetric NATs, based on a changing of a source port, while we explore options with a constant source port.
-
-In our work we study the situation where two parties -- $A, B$ -- try to establish a direct connection, both of them are behind 
-a symmetric NAT (possibly a cascade of such NATs) with a port sensitive, predictable allocation function. Other scenarios with
-predictable allocation function are quite easy to solve and covered in previous research.
+In our work we study the situation where two parties, $A$ and $B$, try to establish a direct connection, both of them are behind 
+a symmetric NAT (possibly a cascade) with a port sensitive, predictable allocation function. 
+%Other scenarios with predictable allocation function are quite easy to solve and covered in previous research.
 
@@
  is a Socket. $\mathbb{IP}$ is the set of all possible IP addresses and $\mathbb{PORT}$ 
-is the set of all possible ports, basically $\mathbb{PORT} = \{1,2,\dots,65535\}$.
+is the set of all possible ports. %, basically $\mathbb{PORT} = \{1,2,\dots,65535\}$.
 % \begin{align*}
@@
 
-\paragraph{NAT and communication.}
+% \paragraph{NAT and communication.}
+\subsection{NAT and communication}
 NAT connects two networks, denoted as {\em internal} and {\em external}, connected to an internal and an external NAT interface, respectively.
@@
 respectively, where $\mathbb{IP}_{in} \cap \mathbb{IP}_{ex} = \emptyset$.
-If it is not stated otherwise the external NAT address\footnote{We assume NAT has only one external address for simplicity.} is denoted as 
+Unless it is not stated otherwise the external NAT address\footnote{We assume NAT has only one external address for simplicity.} is denoted as 
 $\text{IP\textsubscript{nat}} \in \mathbb{IP}_{ex}$.
@@
 \subsection{NAT classes}\label{sec:nat}
-NAT is classified w.r.t. the \emph{filtering rule} and the \emph{allocation rule}, which state what happens after receiving
+NAT is classified with respect to a \emph{filtering rule} and a \emph{allocation rule}, which state what happens after receiving
 packet from an external and an internal network, respectively.
@@
 
-\paragraph{Port Restricted Cone NAT} ~\\
+% \paragraph{Port Restricted Cone NAT} ~\\
+\par\smallskip
+\noindent\textbf{Port Restricted Cone NAT.} 
 Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex} \times \mathbb{PORT}\} \times \mathbb{PORT}$ 
@@
 
-\paragraph{Symmetric NAT} ~\\
+% \paragraph{Symmetric NAT} ~\\
+\par\smallskip
+\noindent\textbf{Symmetric NAT.} 
 Allocation table $\mathbb{AT}$ and filtering rule are the same as in the Port Restricted Cone NAT. The difference lies in the way 
@@
 If the chosen port is already taken, AF iterates until a free one is found.
-
 AF can have different \emph{sensitivity} determining the cases where new allocation is created or the previous one is used instead.
 
-\emph{Address sensitive allocation function} creates new allocation if a destination address differs from the existing allocation 
+\emph{Address sensitive allocation function} creates a~new allocation if a destination address differs from the existing allocation 
 from the same source:
@@
 
-\emph{Port sensitive allocation function} creates new allocation if a destination address and a port differs from the existing
+\emph{Port sensitive allocation function} creates a new allocation if a destination address and a port differs from the existing
 allocations from the same source:
@@
 % \end{align*}
-
 % \section{Related work II}
@@
 All algorithms in our paper make use of the same principle, \emph{UDP hole punching}. 
-Internet hosts cannot send packets to the internal host directly. 
+A public internet host cannot send packets to the private host behind NAT/firewall directly. 
 Communication has to be initiated from the inside\footnote{Not taking
-pre-set port forwarding, DMZ, uPnP, etc. into consideration. }, which is problematic if two nodes behind 
-NAT want to communicate.
+pre-set port forwarding, DMZ, uPnP, etc. into consideration. }, what is problematic if both
+communicating sides are private.
 
@@
 A and B are behind a public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$,
-respectively. By default an internal host does not know its external address and NAT type, thus a technique like 
+respectively. By default an internal host does not know its external address or NAT type, thus a technique like 
 STUN\footnote{Core idea: publicly available service running on 2 IP addresses and 2 ports 
@@
 that is likely to be allocated for a new connection $\text{PORT}^{\text{A}}_{\text{nat}}$ for A and $\text{PORT}^{\text{B}}_{\text{nat}}$
-for B. The period between determining this information, sending to the other party and starting the algorithm is
+for B. The period between determining this information, sending it to the other party and starting the algorithm is
 denoted as an \emph{initialization phase} or a \emph{silent period}\ignore{\footnote{Since the algorithm itself does nothing.}} of the NAT traversal algorithm.
@@
 In order to evaluate algorithms in multiple environments with different parameters (workload, number of clients, etc.)
-we decided to model network behavior with a stochastic process. In particular, we chose a Poisson process, $\{N(t), t\geq0\}$ 
+we decided to model network behavior with a stochastic process. In particular, we chose a \emph{Poisson process}, $\{N(t), t\geq0\}$ 
 that is used in queueing theory to model arrivals of requests to the server queue~\citep{Nelson:1995:PSP:207382}.
@@
 it is important to model how the NAT internal state changes over time, i.e., how many external ports 
-have been allocated in a given time interval. Thus the workload is an important parameter for predicting the allocated port.
+were allocated in a given time interval. Thus the workload is an important parameter for predicting the allocated port.
 
-We model the internal network connected to NAT w.r.t. newly created connections (new port allocation) as a 
-time homogeneous Poisson process $Po(\lambda)$. By setting $\lambda$ we can model 
-different workloads of the network and evaluate algorithms w.r.t. various parameters, e.g., 
+We model the internal network connected to NAT with respect to newly created connections (a new port allocation) as a 
+\emph{time homogeneous Poisson process} $Po(\lambda)$. Value of $\lambda$ models
+different workloads of the network. This way we can evaluate algorithms with respect to various parameters, e.g., 
 probability of establishing a connection.
@@
 \begin{mydef}
-$N(t)$ is a random variable of a number of a new port allocations made in time 
-interval~$[0,t]$ where $N(t) \sim Po(\lambda t)$. \\
-
+$N(t)$ is a random variable of a number of new port allocations made in time 
+interval~$[0,t]$ where $N(t) \sim Po(\lambda t)$.
 \begin{center}                                                     
@@
 
-Moreover, we assume an incremental NAT with $\Delta=1$ on both sides. If we sample the NAT 
+Moreover, we assume an incremental NAT with $\Delta=1$ (without loss of generality) on both sides. If we sample the NAT 
 state\footnote{STUN requests to determine current port number, it creates a new NAT allocation record.}
@@
 
-\paragraph{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected within this time interval.} is assumed to be 3~minutes, 
-which is a typical setting in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~s on average. 
+% \paragraph{Port pool exhaustion.} 
+\par\smallskip
+\noindent\textbf{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected within this time interval.} is assumed to be 3~minutes, 
+what is a typical setting in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~sec on average. 
 If it is higher then NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$.} quickly.
@@
 a connection (i.e., success rate) it proceeds in discrete \emph{steps}. In each step it 
-tries to open a new hole in the local NAT and/or to predict the corresponding port allocated in the destination
+tries to open a new hole in the local NAT and to predict the corresponding port allocated in the destination
 NAT.
@@
 \begin{mydef}
-Symmetric NAT traversal algorithm is a mapping 
+The symmetric NAT traversal algorithm is a mapping 
 $\mathcal{A}: \mathcal{U} \times \mathcal{S} \times \mathcal{M} \rightarrow \mathcal{P}_{src} \times \mathcal{P}_{dst}$, 
-where \\
+where 
 \begin{compactitem}
@@
 parameters and additional information obtained in initial phase). 
-In our setting $\mathcal{M} = \{ (\lambda, t_s) | \lambda~>~0, t\geq~0 \}$. Value $t_s$ denotes the length of the silent period. %For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
+In our setting $\mathcal{M} = \{ (\lambda, t_s) $ $| \lambda~>~0, t\geq~0 \}$. Value $t_s$ denotes the length of the silent period. %For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
 \item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
@@
 \end{mydef}
-In case $\mathcal{M}$ is not not important for the algorithm it can be ignored in the notation.
+If $\mathcal{M}$ is not important for the algorithm it can be ommitted in the notation.
 %If the third parameter is not important for the algorithm it can be ignored in the notation.
@@
 Prior algorithm run it is needed to 
-determine the NAT state (for UDP hole punching), this is assumed to be done by some technique like STUN. For simplicity, we assume that port numbers
+determine the NAT state (for UDP hole punching), this is assumed to be done by a technique like STUN. For simplicity, we assume that port numbers
 start from $0$. The length of the silent period is also known to the algorithm prior execution in order to estimate the current state of 
@@
 
-\paragraph{Notation.}
-Line $A(0,1) = (2,3)$ means party $0$ in step $1$ sent packet from source port $2$ to destination port $3$
-of the party $1$.
+% \paragraph{Notation.}
+\par\smallskip
+\noindent\textbf{Notation.} 
+For instance, line $A(0,1) = (2,3)$ means that party $0$ in step $1$ sent packet from source port $2$ to destination port $3$
+of the party~$1$.
 
-\paragraph{Time.} 
+% \paragraph{Time.} 
+\par\smallskip
+\noindent\textbf{Time.} 
 %The algorithm proceeds in discrete time steps. 
@@
 
-%\paragraph{Party change.} 
-%We define unary operation on communicating parties. 
-%\[
-%a \in \mathcal{U},  \neg a = \begin{dcases*}
-%         0 & when $a=1$\\
-%         1 & when $a=0$
-%        \end{dcases*}
-%\]
-
-\paragraph{Algorithm log \& successful run.} IP addresses of the hosts and NAT are known and fixed prior to the algorithm run, and are thus omitted from the notation for simplicity.
+% \paragraph{Algorithm log \& successful run.} 
+\par\smallskip
+\noindent\textbf{Algorithm log \& successful run.} 
+IP addresses of the hosts and NAT are known and fixed prior to the algorithm run, and are thus omitted from the notation for simplicity.
 For the following text we assume the maximum number of steps of the algorithm $s_{max}=1000$.
@@
 \begin{compactitem}
- \item $(p_1, p_2) = \mathcal{A}(0, s, (\lambda, t_s)) $ are generated by the algorithm $\mathcal{A}$ for some 
+ \item $(p_1, p_2) = \mathcal{A}(0, s, (\lambda, t_s)) $ are generated by the algorithm $\mathcal{A}$ for  
        parameters $(\lambda, t_s)$ in some step $s$ of the algorithm $\mathcal{A}$;
@@
 \end{mydef}
-The task is to find an algorithm $\mathcal{A}$ such that the probability $MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}})$
+The task is to find an algorithm $\mathcal{A}$ such that the probability that $MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}})$
 holds is maximized for minimal number of steps needed.
@@
 \begin{mydef}
-Baby-step giant-step algorithm
+Baby-step, giant-step
 \[
@@
 the expected value is also the most probable value and the algorithm uses maximum likelihood approach.
-\newpage
+% \newpage
 \subsection{Algorithm IV -- Poisson sampling}
@@
 \end{mydef}
+
 %
+
 \subsection{Silent period in algorithms} If the length of the silent period is too big then
@@
 
+% \vspace*{-8mm}
+%\parskip=0pt
+% \titlespacing{\subsection}{0pt}{*0}{*0}
 \section{Evaluation}
 
-\paragraph{Comparison of Algorithms I and II.} 
-Algorithm I has the following benefits over II:
+\subsection{Comparison of Algorithms I and II} 
+Algorithm~I has the following benefits over~II:
 \begin{compactitem}
- \item Works with high success rate for $\lambda \in [0, 0.035] \sim 35$ new connections in 1~sec on average.
+ \item Works with high success rate for $\lambda \in [0, 0.035] \sim 35$ new connections in 1~second on average.
  \item Can be stopped as soon as the connection is established, overhead is proportional to~$\lambda$.
@@
 it\footnote{NAT is abused to multiplex multiple connections to one source port.} compared to Alg. II.
-Due to this fact a practical implementation is simple and lightweight w.r.t. required system resources.
+Due to this fact a practical implementation is simple and lightweight with respect to required system resources.
  \item Is $\lambda$-invariant. Provided network load is low and no additional measurements are needed.
@@
 
-\paragraph{Simulation.}
+% \paragraph{Simulation.}
+\subsection{Simulation}
 
-% \begin{wrapfigure}{r}{0.5\textwidth}
-%  \begin{figure}
-\begin{wrapfigure}{r}{0.5\textwidth}%
-% \wrapfigure{r}{.5\linewidth}{
-     \centering{
-    \begin{subfigure}{0.5\textwidth} 
-	{\scalebox{0.43}{\includegraphics[trim=40 0 0 0]{alg_1.pdf}}}
+\begin{figure}%[H]
+{\scalebox{0.48}{\includegraphics[trim=40 0 0 0]{alg_1.pdf}}}
 \caption{Algorithm evaluation in a simulation, $T=10~ms$. Number of a simulation rounds for one parameter setting is 1000.
  A, B, C correspond to the Algorithm I, II and III respectively.
-	D, E correspond to the Algorithm IV with optimized $C(\lambda T)$ by an exhaustive search in case D, 
-	in case E is then$C(\lambda T)$ is approximated by a polynomial.}
+ D, E correspond to the Algorithm IV with optimized $C(\lambda T)$ by an exhaustive search in D case, 
+ in E case $C(\lambda T)$ is approximated by a polynomial.}
 \label{fig:alg_eval}
-     \end{subfigure}
-%     \end{figure}
+\end{figure}
+
+\begin{figure}%[H]
+{\scalebox{0.48}{\includegraphics[trim=40 0 0 0]{alg_steps.pdf}}}
+\caption{Steps of the algorithms needed to success on average with respect to $\lambda$. Algorithms denoted in a same way as in figure \ref{fig:alg_eval}.}
+\label{fig:alg_steps}
+\end{figure}
+
+% % \begin{wrapfigure}{r}{0.5\textwidth}
+% %  \begin{figure}
+% \begin{wrapfigure}{r}{0.5\textwidth}%
+% % \wrapfigure{r}{.5\linewidth}{
+%      \centering{
+%     \begin{subfigure}{0.5\textwidth} 
+% 	{\scalebox{0.43}{\includegraphics[trim=40 0 0 0]{alg_1.pdf}}}
+% 	\caption{Algorithm evaluation in a simulation, $T=10~ms$. Number of a simulation rounds for one parameter setting is 1000.
+% 	A, B, C correspond to the Algorithm I, II and III respectively.
+% 	D, E correspond to the Algorithm IV with optimized $C(\lambda T)$ by an exhaustive search in case D, 
+% 	in case E is then$C(\lambda T)$ is approximated by a polynomial.}
+% 	\label{fig:alg_eval}
+%      \end{subfigure}
+% %     \end{figure}
+% % \end{wrapfigure}
+% % \begin{wrapfigure}{r}{0.5\textwidth}    
+%      \begin{subfigure}{0.5\textwidth} %\begin{figure}[H]
+% 	{\scalebox{0.43}{\includegraphics[trim=40 0 0 0]{alg_steps.pdf}}}
+% 	\caption{Steps of the algorithms needed to success on average with respect to $\lambda$. 
+% 	Algorithms denoted in a same way as in figure \ref{fig:alg_eval}.}
+% 	\label{fig:alg_steps}
+%      \end{subfigure}}%\end{figure}
+%     \vspace{-30pt}
+% %  \end{figure}
 % \end{wrapfigure}
-% \begin{wrapfigure}{r}{0.5\textwidth}    
-     \begin{subfigure}{0.5\textwidth} %\begin{figure}[H]
-	{\scalebox{0.43}{\includegraphics[trim=40 0 0 0]{alg_steps.pdf}}}
-	\caption{Steps of the algorithms needed to success on average w.r.t. $\lambda$. 
-	Algorithms denoted in a same way as in figure \ref{fig:alg_eval}.}
-	\label{fig:alg_steps}
-     \end{subfigure}}%\end{figure}
-    \vspace{-30pt}
-%  \end{figure}
-\end{wrapfigure}
 
-Figure \ref{fig:alg_eval} shows algorithms success rate w.r.t. $\lambda$. 
+Figure \ref{fig:alg_eval} shows algorithms success rate with respect to~$\lambda$. 
 Figure \ref{fig:alg_steps} depicts number of steps needed 
-to succeed on average. From figures it is visible Alg. I is more suitable for networks with low 
-workload since has $100\%$ success rate for $\lambda \in [0, 0.035]$. Outside this interval 
+to succeed on average. From figures it is visible that Alg. I is more suitable for networks with low 
+workload since has $100\%$ success rate for $\lambda \in [0, 0.035]$. Outside this interval, 
 Alg. I is not working and another solution has to be applied. The problem comes as new connections are
@@
 Data collected in our university network by NetFlow probes were studied to test the correspondence to the Poisson process
-w.r.t. newly created connections/allocations on NAT. NAT was simulated in our program using
+with respect to newly created connections/allocations on NAT. NAT was simulated in our program using
 NetFlow data to provide network traffic to the NAT. NetFlow data capture 1 hour of a network communication in two different
@@
 
-NAT state (newly opened connections in window $T=10, 50, 100~ms$)
+NAT state, e.i., newly opened connections in window $T=10$, $50$, $100$~ms,
 was sampled multiple times in a row, Poisson and Negative Binomial distributions were fitted using Maximum Likelihood Estimation (MLE)
@@
     \hline
-    Set & T [ms] & Samples & Poisson r. [\%] & NBinom r. [\%] & $\overline{m}_{E[X]}$ &  $V[\overline{m}_{E[X]}]$  & $\overline{m}_{V[X]}$ \\ \hline \hline 
+    Set & T [ms] & Samples & Poisson rej. [\%] & NBinom rej. [\%] & $\overline{m}_{E[X]}$ &  $V[\overline{m}_{E[X]}]$  & $\overline{m}_{V[X]}$ \\ \hline \hline 
     \multirow{7}{*}{03} & \multirow{3}{*}{10}   & 100		& 2.977		& 1.667		& 0.5121        & 0.0228                    & 0.5321 \\ \cline{3-8}
@@
       &      & 500     		& 57.746	& 9.859		& 4.7668	& 0.2876 	& 7.4667 					\\ \hline
-
     \multirow{2}{*}{13}  & \multirow{2}{*}{10}     & 100     & 5.191      & 1.824     & 2.6438       &  0.2695    & 3.0020 \\ \cline{3-8}
@@
     \% of cases rejecting hypothesis about Negative Binomial distribution,
-    sample mean of sample mean of port samples,
-    sample variance of sample mean of port samples,
-    sample mean of variance of port samples.}
+    sample mean of sample mean of port number samples (pns),
+    sample variance of sample mean of pns,
+    sample mean of variance of pns.}
   \label{tab:1}
@@
 
-Regarding data from set 13, during the activity peak during the day, we conclude that if a number of samples 
+Regarding data from set 13, which represents the high peak activity during the day, we conclude that if a number of samples 
 is lower, e.g., 100, the process of creating a new connection on NAT can 
@@
 
-If the number of samples is higher, e.g., 500, the time window of studied network traffic is longer and
+When the number of samples is higher, e.g., 500, the time window of the studied network traffic is longer and
 in a real world scenario the probability that $\lambda$ is constant in time is lower with the increasing length of the time
@@
 data (hypothesis rejected in $2.829\%$ of cases). 
-The open question is how to make use of this model in the NAT traversal.
+%The open question is how to make use of this model in the NAT traversal.
 
@@
 
-Regarding data from set 03, we can conclude that a lower network workload causes more stable mean of sampled ports (MLE for Poisson distribution), thus the Poisson distribution hypothesis is not rejected even in longer time intervals for an estimated $\lambda$.
+Regarding data from set 03, which represents the low peak activity we can conclude that a lower network workload causes more stable mean of sampled ports (MLE for Poisson distribution), thus the Poisson distribution hypothesis is not rejected even in longer time intervals for an estimated $\lambda$.
 
 \section{Discussion and implementation}
-For simplicity, we assumed that $\lambda$ is the same on both sides, the new challenges 
-can arise if we allow different parameters on both sides that we have not studied yet.
+For sake of the paper, we assumed that $\lambda$ is the same on both sides. However, new challenges 
+are likely to arise if we allow for different parameters on both sides, and we have not studied this yet.
 
@@
 we do not have exact analytic explanation. This method is also very sensitive to parameter fluctuations
-and thus not very practical for a real world networks.
+and thus not very practical for real world networks.
 
-In our setting we assume $T=10~ms$ in order to minimize the speed of the evolution of a Poisson process.
+In our setting, we assume $T=10~ms$ in order to minimize the speed of the evolution of a Poisson process.
 Such a short sending interval may cause some packets to get dropped on the path to the destination. Yet the 
@@
 
-Success rate of the algorithm strongly relies on accurate measurements of the network 
-properties the client resides in. They can be measured prior to the connection establishment 
+Success rate of the algorithms strongly relies on accurate measurements of the network 
+properties the client resides in. These properties can be measured prior to the connection establishment 
 from the point the client connects to the network, as is usually done in ICE, and to gather 
@@
 
-%We ignored the cases where there are already some allocations on NAT from older connections
-%on the port interval the NAT traversal algorithm is using.
-
-Another problematic issue comes when this algorithm is followed by multiple host pairs on
-the same network in the same time. They would artificially increase the network workload
-causing it would not follow Poisson distribution rendering our algorithms not usable in
-this case.
+% %We ignored the cases where there are already some allocations on NAT from older connections
+% %on the port interval the NAT traversal algorithm is using.
+% 
+% Another problematic issue comes when this algorithm is followed by multiple host pairs on
+% the same network in the same time. They would artificially increase the network workload
+% causing it would not follow Poisson distribution rendering our algorithms not usable in
+% this case.
 
@@
 The aim of our work was to study the problem of NAT traversal from a statistical point of view,
-and to provide an algorithm (or algorithms) with a high success rate for networks with a low workload that would be easy to 
+and to provide algorithms with a high success rate for networks with a low workload that would be easy to 
 implement and would require little system resources.
@@
 We modeled the network workload as a Poisson process and we verified our approach on real data by hypothesis testing.
-Our testing application was published in public domain at GitHub, and we also validated our results by traversing symmetric 
-carrier-grade NAT of our mobile phone operator.
+Our testing application was published in public domain. We also validated our results by traversing symmetric 
+carrier-grade NAT of our mobile phone provider.
 
-\appendix
-\section{NAT classes}\label{sec:classes_appendix}
-Definition of a Full cone NAT and Address restricted cone NAT as given in section \ref{sec:nat}.
+\par\smallskip
+\noindent\textbf{Acknowledgement.} 
+This work was supported by the Czech Science Foundation, project GAP202/11/0422. TODO: ACK netflow Vykopall
 
-\paragraph{Full cone NAT.} ~\\
-Allocation table $\mathbb{AT} \subseteq \{ \mathbb{IP}_{in} \times \mathbb{PORT} \} \times \mathbb{PORT}$ holds associations between internal 
-hosts socket and mapped external NAT port. 
-% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
-Filtering rule is:
-\begin{align*}
-& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
-&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}_{in}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
-& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
-\end{align*}
-
-\paragraph{Address restricted cone NAT.} ~\\
-Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex}\} \times \mathbb{PORT}$ 
-holds associations between internal hosts socket, external host IP address and mapped external NAT port. 
-% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
-Filtering rule is:
-\begin{align*}
-& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
-&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
-& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
-\end{align*}
-
-% -----
-% Bibliography
-%\bibliographystyle{unsrtnat}
-\bibliographystyle{plainnat}
+% \renewcommand{\bibfont}{\small}
+% \bibliographystyle{IEEEtran}
+\bibliographystyle{abbrv}
 \bibliography{paper}

### 2014-07-16 NAT_paper_ACM_short.tex
added github link
almost there

@@
 \begin{figure}%[H]
-{\scalebox{0.48}{\includegraphics[trim=40 0 0 0]{alg_1.pdf}}}
+{\scalebox{0.48}{\includegraphics[trim=40 20 0 0]{alg_1.pdf}}}
 \caption{Algorithm evaluation in a simulation, $T=10~ms$. Number of a simulation rounds for one parameter setting is 1000.
@@
 \begin{figure}%[H]
-{\scalebox{0.48}{\includegraphics[trim=40 0 0 0]{alg_steps.pdf}}}
+{\scalebox{0.48}{\includegraphics[trim=40 20 0 0]{alg_steps.pdf}}}
 \caption{Steps of the algorithms needed to success on average with respect to $\lambda$. Algorithms denoted in a same way as in figure \ref{fig:alg_eval}.}
@@
 may be helpful for another researchers interested in this area so we published it in public domain
-hosted on GitHub [URL:HERE].
+hosted on GitHub: \url{https://github.com/ph4r05/NATSimTools}.
 
@@
 
+We believe that by removing unnecessary intermediate nodes from
+the communication path we restrict the attack and failure surfaces and increase the communication reliability.
+Communication paths get less complex, transparency increases and this all support a more natural way of the 
+interconnecting nodes on P2P basis.
+
 \par\smallskip
 \noindent\textbf{Acknowledgement.} 
-This work was supported by the Czech Science Foundation, project GAP202/11/0422. TODO: ACK netflow Vykopall
+This work was supported by the Czech Science Foundation, project GAP202/11/0422. The netflow data was provided
+by the CSIRT team of the Masaryk University, Brno, Czech Republic TODO: FIX.
 
